---
title: "HB Temporal C Synthesis"
author: "Sophie von Fromm"
date: "`r Sys.Date()`"
output: html_document
---
#Setup
```{r setup, include=FALSE}
rm(list = ls()) #clear R environment

# Check for and install required packages and load libraries
for (package in c( 'tidyverse', 'here')) {    #can add more packages to this vector
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}

knitr::opts_knit$set(root.dir = here())

```

#TITLE FROM EDI: Continuous precipitation and stream chemistry data, Hubbard Brook Ecosystem Study, 1963 – ongoing 
DESCRIBE DATA: Long-term monitoring of Stream and Precipitation Chemistry. Focusing on precipitation pH, DOC
specific conductivity, and NO3.

#Dealing with NA codes
DESCRIBE the NA system for your dataset: 

NA, Data missing, or not taken at this time

#Access data from EDI
Note where you need to paste in the url of your dataset
Note where you need to add in NA codes that you described above for your dataset

```{r access data from EDI}
options(HTTPUserAgent="EDI_CodeGen")

inUrl1  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/208/14/f148543b93982fd52f7089d2f2f51e85"
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")

dt1 <- read.csv(infile1, stringsAsFactors=F, na.strings = c("NA")) 
unlink(infile1)

rm(inUrl1, infile1)
```

#Clean up your data
DESCRIBE how you clean up your data (below is example from Nina's work on the stream chemistry data).
YOU will have to convert the Date field properly using as.Date. If date is stored in a YYYY-MM-DD format, no need to specify the format. If not, scroll down to bottom of this page to see how to describe various formats: https://www.r-bloggers.com/2013/08/date-formats-in-r/

Removed samples flagged as duplicates (filter != "Dup)

Removed sites == "S". From 1963-1969, South facing precipitation samples were 
chosen from the “cleanest” precipitation data and reported as site=S

#Make the changes.


```{r data cleaning}

summary(dt1) 

#Data looks good at a first glimpse

dt1 %>% 
  count(duplicate)

dt2 <- dt1 %>% 
  dplyr::filter(duplicate != "Dup" | is.na(duplicate)) %>% 
  dplyr::filter(site != "S" | is.na(site)) %>% 
  dplyr::select(site, date, uniqueID, fieldCode, waterYr, duplicate,
                pH, DOC, spCond, NO3, SO4)

summary(dt2)

dt2 %>% 
  count(site)

dt2 %>% 
  count(duplicate)


```

#Filter data to the window of time we are interested in
```{r}
# W6_streamconc <- W6_streamconc %>% 
#   filter(Date > "1980-01-01" & Date< "2022-12-31")
```

##STOPPED HERE##

#Summarize your data by month and year for graphing
DESCRIBE timestep(s) of your data here: Monthly
Grouped by waterYr for now
Averaged analyte for time-step; assuming it is corrected for volume (should be double-checked again)
Averaged across all sites (rain gauges)

```{r}

atm_WaterYear <- dt2 %>%
  group_by(waterYr) %>% 
  summarise(across(
      where(is.numeric), 
      list(mean = ~ mean(.x, na.rm = TRUE), 
           sd = ~ sd(.x, na.rm = TRUE),
           se   = ~ sd(.x, na.rm = TRUE) / sqrt(sum(!is.na(.x))))),
    .groups = "drop"
  )

summary(atm_WaterYear)
```

#Graph your analytes by year in ggplot
# y-scales are on different scales
```{r}
atm_WaterYear_pH <- atm_WaterYear %>% 
  drop_na(pH_mean) %>% 
  ggplot(aes(x = waterYr, y = pH_mean)) +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = pH_mean - pH_se, ymax = pH_mean + pH_se)) +
  geom_line() +
  labs(
    x = "Water year",
    y = "pH in precipitation",
    title = "Water year pH in precipitation (mean + SE) ") +
  theme_minimal()

atm_WaterYear_DOC <- atm_WaterYear %>% 
  drop_na(DOC_mean) %>%
  ggplot(aes(x = waterYr, y = DOC_mean)) +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = DOC_mean - DOC_se, ymax = DOC_mean + DOC_se)) +
  geom_line() +
  labs(
    x = "Water year",
    y = "DOC (mg/L) in precipitation",
    title = "Water year DOC (mg/L) in precipitation (mean + SE) ") +
  theme_minimal()

atm_WaterYear_spCond <- atm_WaterYear %>% 
  drop_na(spCond_mean) %>%
  ggplot(aes(x = waterYr, y = spCond_mean)) +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = spCond_mean - spCond_se, ymax = spCond_mean + spCond_se)) +
  geom_line() +
  labs(
    x = "Water year",
    y = "Specific conductivity (uS/cm) in precipitation",
    title = "Water year specific conductivity (uS/cm) in precipitation (mean + SE) ") +
  theme_minimal()

atm_WaterYear_NO3 <- atm_WaterYear %>% 
  drop_na(NO3_mean) %>%
  ggplot(aes(x = waterYr, y = NO3_mean)) +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = NO3_mean - NO3_se, ymax = NO3_mean + NO3_se)) +
  geom_line() +
  labs(
    x = "Water year",
    y = "NO3 (mg/L) in precipitation",
    title = "Water year nitrate (mg/L) in precipitation (mean + SE) ") +
  theme_minimal()

atm_WaterYear_SO4 <- atm_WaterYear %>% 
  drop_na(SO4_mean) %>%
  ggplot(aes(x = waterYr, y = SO4_mean)) +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = SO4_mean - SO4_se, ymax = SO4_mean + SO4_se)) +
  geom_line() +
  labs(
    x = "Water year",
    y = "SO4 (mg/L) in precipitation",
    title = "Water year sulfate (mg/L) in precipitation (mean + SE) ") +
  theme_minimal()

```

#Save all your plots to the graph folder
```{r}
plots <- list(atm_WaterYear_pH, atm_WaterYear_DOC, atm_WaterYear_spCond, 
              atm_WaterYear_NO3, atm_WaterYear_SO4) 
names <- c("atm_WaterYear_pH", "atm_WaterYear_DOC", "atm_WaterYear_spCond", 
           "atm_WaterYear_NO3", "atm_WaterYear_SO4") 

for (i in seq_along(plots)) {
  ggsave(
    filename = here::here("Graphs/Atmosphere", paste0(names[i], ".png")),
    plot = plots[[i]],
    width = 8,
    height = 5,
    dpi = 300
  )
}
```

LASTLY push your changes to our github repo!!