---
title: "HB Temporal Carbon - W1 Lysimeter Data"
author: "Angela Possinger"
date: "`r Sys.Date()`"
output: html_document
---
#Setup
```{r setup, include=FALSE}
rm(list = ls()) #clear R environment

# Check for and install required packages and load libraries
for (package in c( 'tidyverse', 'here')) {    #can add more packages to this vector
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}

knitr::opts_knit$set(root.dir = here())

```

#TITLE FROM EDI: Chemistry of freely-draining soil solutions at the Hubbard Brook Experimental Forest, Watershed 1, 1996 - present
DESCRIBE DATA: W1 lysimeter soil solution chemistry data, approximately monthly since 1996 
Focus variables: pH, ANC, Al, Si, NH4, NO3, DOC, DIC, TN

#Dealing with NA codes
Here are the descriptions of missing value codes from the metadata. The codes are the same for all variables, so that makes things easy. In this script, I code -999.99, -888.88 and -666.66 as NA when I load the data. Then I deal with -777.77 separately. 

-999.99, Data missing, or not taken at this time
-888.88, Sample volume too small for analysis
-666.66, Result Not Reported: Failed Internal QA/QC Review
-777.77, Analyte concentration below limit of detection

#Access data from EDI
Note where you need to paste in the url of your dataset
Note where you need to add in NA codes that you described above for your dataset

```{r access data from EDI}
options(HTTPUserAgent="EDI_CodeGen")

inUrl1  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/138/14/429a88e19df84e6b1ade16a1ab434f62" #paste url here
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")

dt1 <-read.csv(infile1, stringsAsFactors=F, na.strings = c("-999.99", "-888.88", "-666.66")) #add in NA codes here!
unlink(infile1)

rm(inUrl1, infile1)
```

#Clean up your data
DESCRIBE how you clean up your data (below is example from Nina's work on the stream chemistry data).
YOU will have to convert the Date field properly using as.Date. If date is stored in a YYYY-MM-DD format, no need to specify the format. If not, scroll down to bottom of this page to see how to describe various formats: https://www.r-bloggers.com/2013/08/date-formats-in-r/

We will need to substitute half the detection limit for these data to avoid '-777.77' being treated as a numeric value. Note that if were to use the automatically generated EDI code, '-777.77' would become NA. Also, some values in the data table are below the stated detection limit for the analyte even though they do not have a -777.77 code, so with this step we are only dealing with the -777.77 but not making any decision about other numeric values that are below the limit. 

Calculated variables such as ANC are not easily estimated by detection limits; -777.77 values for ANC will be coded as NA. 

pH doesn't have a logical detection limit; -777.77 values for pH will be coded as NA. 

#Description of detection limits from the methods metadata: "Current detection limits for the Driscoll laboratory: 
Na (0.015 umol/L), K (0.12 umol/L), Ca (0.030 umol/L), 
Mg (0.0032 umol/L), Alt (0.0026 umol/L), Alm (1.3 umol/L), 
Alo (1.3 umol/L), NH4 (0.55 umol/L), H4SiO4 (3 umol/L), 
Cl (0.55 umol/L), NO3 (0.85 umol/L), SO4 (0.53 umol/L), 
F (0.097 umol/L), DIC (20 umol/L), DOC (20 umol/L), 
TN (0.81 umol/L). It is suggested to dataset users to use a value of one half the detection limit when a value for a solute is less than the detection limit.

#Make the changes.
As you will see below, in this example some judgement calls had to be made

```{r clean up data}

#Remove "." from column names
names(dt1) <- gsub("\\.", "", names(dt1))

#Make a 'look-up' list of values to give measurements below detection limit
detectionLimits_Driscoll <- list(
  Na = 0.015/2,
  K = 0.12/2,
  Ca2 = 0.030/2,
  Mg2 = 0.0032/2,
  Alt = 0.0026/2,
  Altm = 1.3/2,
  Alom = 1.3/2,
  NH4 = 0.55/2,
  H4SiO4 = 3/2,
  NO3 = 0.85/2,
  SO4 = 0.53/2,
  F = 0.097/2,
  DIC = 20/2,
  DOC = 20/2,
  TN = 0.81/2,
  Cl = 0.55/2, 
  ANC = NA, 
  pH = NA
)

W1_lysimconc <- dt1 %>%
  mutate(Date = as.Date(Date)) %>% #convert to Date object
  mutate(across(  #replace -777.77 (values below the detection limit) with half the limit
    .cols = names(detectionLimits_Driscoll),
    .fns = ~ ifelse(.x == -777.77, detectionLimits_Driscoll[[cur_column()]], .x)))

#take a look to make sure the data is behaving how you want it to
summary(W1_lysimconc) 

#Extreme low values: pH, ANC, Al, NH4, NO3, DOC, DIC, TN
#pH
unique(W1_lysimconc$pH[which(W1_lysimconc$pH < 0)]) 
#They are:
#[1] -1000
#take a closer look
#View(W1_lysimconc[which(W1_lysimconc$pH < 0),])
#These rows are also missing data for all other parameters

#ANC
#Distribution of ANC 
hist(W1_lysimconc$ANC)
unique(W1_lysimconc$ANC[which(W1_lysimconc$ANC > 400)]) 
#They are:
#[1] 453.10 783.90 415.00 445.94 509.41 428.06 414.95
#take a closer look
#View(W1_lysimconc[which(W1_lysimconc$ANC > 400),])
#These values are consistent with relatively high pH - do not remove

#Alt
unique(W1_lysimconc$Alt[which(W1_lysimconc$Alt < -100)]) 
#They are:
#[1] -336.4
#take a closer look
#View(W1_lysimconc[which(W1_lysimconc$Alt < -100),])
#Base cations also erroneously low in this sample

#Altm
unique(W1_lysimconc$Altm[which(W1_lysimconc$Altm < -100)]) 
#They are:
#  [1] -774.23 -772.26 -776.22 -775.97 -775.71 -768.30 -773.34 -775.31
#  [9] -775.69 -775.27 -776.54 -774.32 -772.12 -772.00 -766.24 -769.81
# [17] -773.13 -772.53 -774.03 -772.16 -770.01 -774.98 -768.14 -770.10
# [25] -768.52 -764.14 -770.19 -773.10 -771.73 -771.55 -775.17 -774.36
# [33] -768.90 -770.14 -772.73 -771.66 -776.26 -772.50 -772.94 -776.41
# [41] -774.44 -775.38 -766.66 -770.46 -771.70 -770.45 -775.90 -776.01
# [49] -771.57 -776.38 -776.28 -767.80 -774.21 -772.17 -775.60 -771.23
# [57] -774.11 -776.19 -772.74 -776.32 -775.09 -774.81 -773.40 -768.73
# [65] -772.88 -774.27 -385.06
#take a closer look
#View(W1_lysimconc[which(W1_lysimconc$Altm < -100),])
#All low only for Altm - other columns kept 

#Alom
unique(W1_lysimconc$Alom[which(W1_lysimconc$Alom < -10)]) 
#They are:
#None

#NH4
unique(W1_lysimconc$NH4[which(W1_lysimconc$NH4 < -100)]) 
#They are:
#None

#NO3
unique(W1_lysimconc$NO3[which(W1_lysimconc$NO3 < -100)]) 
#They are:
#None

#DOC
unique(W1_lysimconc$DOC[which(W1_lysimconc$DOC < -100)]) 
#They are:
#None

#DIC
unique(W1_lysimconc$DIC[which(W1_lysimconc$DIC < -100)]) 
#They are:
#None

#TN
unique(W1_lysimconc$TN[which(W1_lysimconc$TN < -10)]) 
#They are:
#57.65
#take a closer look
#View(W1_lysimconc[which(W1_lysimconc$TN < -10),])
#Low only for TN - other columns kept 

#Change the pH, Al, etc. with values defined above to NA and then remove rows from the df with no chemistry data for any analyte:
W1_lysimconc <- W1_lysimconc %>%
  mutate(pH = ifelse(pH < 0, NA, pH)) %>%
  mutate(Alt = ifelse(Alt < -100, NA, Alt)) %>%
  mutate(Altm = ifelse(Altm < -100, NA, Altm)) %>%
  mutate(TN = ifelse(TN < -10, NA, TN)) %>%
  filter(!if_all(pH:TN, is.na))

#Convert the character columns to factors (needed for month and year function below)
W1_lysimconc <- W1_lysimconc %>%
  mutate(across(c(WS, Site, Horizon), as.factor))

#Add column for "organic" and "mineral" to split horizon types
W1_lysimconc <- W1_lysimconc %>%
  mutate(Horizon_type = ifelse(
    Horizon %in% c("Oe", "Oae", "Oa"),
    "Organic",
    "Mineral"
  ))


#clean up
rm(dt1) #remove the original uncleaned data file

```

#Filter data to the window of time we are interested in
For now, keep it to its native time frame
# ```{r}
# W6_streamconc <- W6_streamconc %>% 
#   filter(Date > "1980-01-01" & Date< "2022-12-31")
# ```

#Summarize your data by month and year for graphing
DESCRIBE timestep(s) of your data here: Monthly
In the example below we are averaging over sites for each month
```{r}
W1_lysimconc_month <- W1_lysimconc %>%
  mutate(month = floor_date(Date, "month")) %>% #by month, converts all dates to 1st of month
  group_by(month, WS, Horizon, Horizon_type) %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)),
            .groups = "drop")

W1_lysimconc_year <- W1_lysimconc %>%
  mutate(year = floor_date(Date, "year")) %>% #by year, converts all dates to 1st of year
  group_by(year, WS, Horizon, Horizon_type) %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)),
            .groups = "drop")

#save these?? Right now no, and there is no data folder in the project
#write.csv(data_name, here("data", "data_name.csv"))
```

#Graph your analytes by month and year in ggplot
```{r}
#pH#### 
#Month
#Organic
pH_month_organic <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Organic"), aes(x = month, y = pH)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution pH",
    title = "Organic Horizon Monthly Mean Soil Solution pH Over Time") +
  theme_minimal()

#Mineral
pH_month_mineral <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Mineral"), aes(x = month, y = pH)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution pH",
    title = "Mineral Horizon Monthly Mean Soil Solution pH Over Time") +
  theme_minimal()

#Year 
#Organic
pH_year_organic <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Organic"), aes(x = year, y = pH)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution pH",
    title = "Organic Horizon Yearly Mean Soil Solution pH Over Time") +
  theme_minimal()

#Mineral
pH_year_mineral <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Mineral"), aes(x = year, y = pH)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution pH",
    title = "Mineral Horizon Yearly Mean Soil Solution pH Over Time") +
  theme_minimal()

#ANC#### 
#Month
#Organic
ANC_month_organic <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Organic"), aes(x = month, y = ANC)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution ANC",
    title = "Organic Horizon Monthly Mean Soil Solution ANC Over Time") +
  theme_minimal()

#Mineral
ANC_month_mineral <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Mineral"), aes(x = month, y = ANC)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution ANC",
    title = "Mineral Horizon Monthly Mean Soil Solution ANC Over Time") +
  theme_minimal()

#Year 
#Organic
ANC_year_organic <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Organic"), aes(x = year, y = ANC)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution ANC",
    title = "Organic Horizon Yearly Mean Soil Solution ANC Over Time") +
  theme_minimal()

#Mineral
ANC_year_mineral <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Mineral"), aes(x = year, y = ANC)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution ANC",
    title = "Mineral Horizon Yearly Mean Soil Solution ANC Over Time") +
  theme_minimal()

#Alt#### 
#Month
#Organic
Alt_month_organic <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Organic"), aes(x = month, y = Alt)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution Alt",
    title = "Organic Horizon Monthly Mean Soil Solution Alt Over Time") +
  theme_minimal()

#Mineral
Alt_month_mineral <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Mineral"), aes(x = month, y = Alt)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution Alt",
    title = "Mineral Horizon Monthly Mean Soil Solution Alt Over Time") +
  theme_minimal()

#Year 
#Organic
Alt_year_organic <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Organic"), aes(x = year, y = Alt)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution Alt",
    title = "Organic Horizon Yearly Mean Soil Solution Alt Over Time") +
  theme_minimal()

#Mineral
Alt_year_mineral <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Mineral"), aes(x = year, y = Alt)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution Alt",
    title = "Mineral Horizon Yearly Mean Soil Solution Alt Over Time") +
  theme_minimal()

#Altm#### 
#Month
#Organic
Altm_month_organic <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Organic"), aes(x = month, y = Altm)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution Altm",
    title = "Organic Horizon Monthly Mean Soil Solution Altm Over Time") +
  theme_minimal()

#Mineral
Altm_month_mineral <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Mineral"), aes(x = month, y = Altm)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution Altm",
    title = "Mineral Horizon Monthly Mean Soil Solution Altm Over Time") +
  theme_minimal()

#Year 
#Organic
Altm_year_organic <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Organic"), aes(x = year, y = Altm)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution Altm",
    title = "Organic Horizon Yearly Mean Soil Solution Altm Over Time") +
  theme_minimal()

#Mineral
Altm_year_mineral <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Mineral"), aes(x = year, y = Altm)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution Altm",
    title = "Mineral Horizon Yearly Mean Soil Solution Altm Over Time") +
  theme_minimal()

#Alom#### 
#Month
#Organic
Alom_month_organic <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Organic"), aes(x = month, y = Alom)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution Alom",
    title = "Organic Horizon Monthly Mean Soil Solution Alom Over Time") +
  theme_minimal()

#Mineral
Alom_month_mineral <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Mineral"), aes(x = month, y = Alom)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution Alom",
    title = "Mineral Horizon Monthly Mean Soil Solution Alom Over Time") +
  theme_minimal()

#Year 
#Organic
Alom_year_organic <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Organic"), aes(x = year, y = Alom)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution Alom",
    title = "Organic Horizon Yearly Mean Soil Solution Alom Over Time") +
  theme_minimal()

#Mineral
Alom_year_mineral <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Mineral"), aes(x = year, y = Alom)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution Alom",
    title = "Mineral Horizon Yearly Mean Soil Solution Alom Over Time") +
  theme_minimal()

#NH4#### 
#Month
#Organic
NH4_month_organic <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Organic"), aes(x = month, y = NH4)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution NH4",
    title = "Organic Horizon Monthly Mean Soil Solution NH4 Over Time") +
  theme_minimal()

#Mineral
NH4_month_mineral <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Mineral"), aes(x = month, y = NH4)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution NH4",
    title = "Mineral Horizon Monthly Mean Soil Solution NH4 Over Time") +
  theme_minimal()

#Year 
#Organic
NH4_year_organic <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Organic"), aes(x = year, y = NH4)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution NH4",
    title = "Organic Horizon Yearly Mean Soil Solution NH4 Over Time") +
  theme_minimal()

#Mineral
NH4_year_mineral <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Mineral"), aes(x = year, y = NH4)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution NH4",
    title = "Mineral Horizon Yearly Mean Soil Solution NH4 Over Time") +
  theme_minimal()

#NO3#### 
#Month
#Organic
NO3_month_organic <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Organic"), aes(x = month, y = NO3)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution NO3",
    title = "Organic Horizon Monthly Mean Soil Solution NO3 Over Time") +
  theme_minimal()

#Mineral
NO3_month_mineral <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Mineral"), aes(x = month, y = NO3)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution NO3",
    title = "Mineral Horizon Monthly Mean Soil Solution NO3 Over Time") +
  theme_minimal()

#Year 
#Organic
NO3_year_organic <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Organic"), aes(x = year, y = NO3)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution NO3",
    title = "Organic Horizon Yearly Mean Soil Solution NO3 Over Time") +
  theme_minimal()

#Mineral
NO3_year_mineral <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Mineral"), aes(x = year, y = NO3)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution NO3",
    title = "Mineral Horizon Yearly Mean Soil Solution NO3 Over Time") +
  theme_minimal()

#DOC#### 
#Month
#Organic
DOC_month_organic <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Organic"), aes(x = month, y = DOC)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution DOC",
    title = "Organic Horizon Monthly Mean Soil Solution DOC Over Time") +
  theme_minimal()

#Mineral
DOC_month_mineral <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Mineral"), aes(x = month, y = DOC)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution DOC",
    title = "Mineral Horizon Monthly Mean Soil Solution DOC Over Time") +
  theme_minimal()

#Year 
#Organic
DOC_year_organic <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Organic"), aes(x = year, y = DOC)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution DOC",
    title = "Organic Horizon Yearly Mean Soil Solution DOC Over Time") +
  theme_minimal()

#Mineral
DOC_year_mineral <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Mineral"), aes(x = year, y = DOC)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution DOC",
    title = "Mineral Horizon Yearly Mean Soil Solution DOC Over Time") +
  theme_minimal()

#DIC#### 
#Month
#Organic
DIC_month_organic <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Organic"), aes(x = month, y = DIC)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution DIC",
    title = "Organic Horizon Monthly Mean Soil Solution DIC Over Time") +
  theme_minimal()

#Mineral
DIC_month_mineral <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Mineral"), aes(x = month, y = DIC)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution DIC",
    title = "Mineral Horizon Monthly Mean Soil Solution DIC Over Time") +
  theme_minimal()

#Year 
#Organic
DIC_year_organic <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Organic"), aes(x = year, y = DIC)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution DIC",
    title = "Organic Horizon Yearly Mean Soil Solution DIC Over Time") +
  theme_minimal()

#Mineral
DIC_year_mineral <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Mineral"), aes(x = year, y = DIC)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution DIC",
    title = "Mineral Horizon Yearly Mean Soil Solution DIC Over Time") +
  theme_minimal()

#TN#### 
#Month
#Organic
TN_month_organic <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Organic"), aes(x = month, y = TN)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution TN",
    title = "Organic Horizon Monthly Mean Soil Solution TN Over Time") +
  theme_minimal()

#Mineral
TN_month_mineral <- ggplot(subset(W1_lysimconc_month, Horizon_type == "Mineral"), aes(x = month, y = TN)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution TN",
    title = "Mineral Horizon Monthly Mean Soil Solution TN Over Time") +
  theme_minimal()

#Year 
#Organic
TN_year_organic <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Organic"), aes(x = year, y = TN)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution TN",
    title = "Organic Horizon Yearly Mean Soil Solution TN Over Time") +
  theme_minimal()

#Mineral
TN_year_mineral <- ggplot(subset(W1_lysimconc_year, Horizon_type == "Mineral"), aes(x = year, y = TN)) +
  geom_point(size = 1) +
  geom_line() +
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution TN",
    title = "Mineral Horizon Yearly Mean Soil Solution TN Over Time") +
  theme_minimal()

```

#Save all your plots to the graph folder
```{r}
plots <- list(pH_month_organic, pH_month_mineral, pH_year_organic, pH_year_mineral, ANC_month_organic, ANC_month_mineral, ANC_year_organic, ANC_year_mineral, Alt_month_organic, Alt_month_mineral, Alt_year_organic, Alt_year_mineral, Altm_month_organic, Altm_month_mineral, Altm_year_organic, Altm_year_mineral, Alom_month_organic, Alom_month_mineral, Alom_year_organic, Alom_year_mineral, NH4_month_organic, NH4_month_mineral, NH4_year_organic, NH4_year_mineral, NO3_month_organic, NO3_month_mineral, NO3_year_organic, NO3_year_mineral, DOC_month_organic, DOC_month_mineral, DOC_year_organic, DOC_year_mineral, DIC_month_organic, DIC_month_mineral, DIC_year_organic, DIC_year_mineral, TN_month_organic, TN_month_mineral, TN_year_organic, TN_year_mineral ) #change to your graph names

names <- c( "pH_month_organic", "pH_month_mineral",
  "pH_year_organic", "pH_year_mineral",
  "ANC_month_organic", "ANC_month_mineral",
  "ANC_year_organic", "ANC_year_mineral",
  "Alt_month_organic", "Alt_month_mineral",
  "Alt_year_organic", "Alt_year_mineral",
  "Altm_month_organic", "Altm_month_mineral",
  "Altm_year_organic", "Altm_year_mineral",
  "Alom_month_organic", "Alom_month_mineral",
  "Alom_year_organic", "Alom_year_mineral",
  "NH4_month_organic", "NH4_month_mineral",
  "NH4_year_organic", "NH4_year_mineral",
  "NO3_month_organic", "NO3_month_mineral",
  "NO3_year_organic", "NO3_year_mineral",
  "DOC_month_organic", "DOC_month_mineral",
  "DOC_year_organic", "DOC_year_mineral",
  "DIC_month_organic", "DIC_month_mineral",
  "DIC_year_organic", "DIC_year_mineral",
  "TN_month_organic", "TN_month_mineral",
  "TN_year_organic", "TN_year_mineral"
) #change to names

for (i in seq_along(plots)) {
  ggsave(
    filename = here::here("Graphs", paste0(names[i], ".png")),
    plot = plots[[i]],
    width = 8,
    height = 10,
    dpi = 300
  )
}
```

LASTLY push your changes to our github repo!!