---
title: "HB Temporal C Synthesis"
author: "Sophie von Fromm"
date: "`r Sys.Date()`"
output: html_document
---
#Setup
```{r setup, include=FALSE}
rm(list = ls()) #clear R environment

# Check for and install required packages and load libraries
for (package in c( 'tidyverse', 'here')) {    #can add more packages to this vector
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}

knitr::opts_knit$set(root.dir = here())

```

#TITLE FROM EDI: Long-term measurements of microbial biomass and activity at the Hubbard Brook Experimental Forest 1994 – ongoing 
DESCRIBE DATA: Long-term monitoring of soil nitrate (NO3-) and ammonium (NH4+) concentrations, microbial biomass carbon (C) and nitrogen (N) content, microbial respiration, potential nitrification and N mineralization rates, pH, and denitrification potential for west of watershed 6 and watershed 1

#Dealing with NA codes
DESCRIBE the NA system for your dataset: 

-999.99, Data missing, or not taken at this time

#Access data from EDI
Note where you need to paste in the url of your dataset
Note where you need to add in NA codes that you described above for your dataset

```{r access data from EDI}
options(HTTPUserAgent="EDI_CodeGen")

inUrl1  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/67/26/a86c0c9ba835e4e09709ed42969c18be"
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")

dt1 <- read.csv(infile1, stringsAsFactors=F, na.strings = c("-9999.99")) 
unlink(infile1)

rm(inUrl1, infile1)
```

#Clean up your data
DESCRIBE how you clean up your data (below is example from Nina's work on the stream chemistry data).
YOU will have to convert the Date field properly using as.Date. If date is stored in a YYYY-MM-DD format, no need to specify the format. If not, scroll down to bottom of this page to see how to describe various formats: https://www.r-bloggers.com/2013/08/date-formats-in-r/

#Make the changes.
As you will see below, in this example some judgement calls had to be made

```{r data cleaning}

summary(dt1) 

#Data looks good at a first glimpse


```

#Filter data to the window of time we are interested in
```{r}
# W6_streamconc <- W6_streamconc %>% 
#   filter(Date > "1980-01-01" & Date< "2022-12-31")
```


#Summarize your data by month and year for graphing
DESCRIBE timestep(s) of your data here: Yearly
In the example below we are averaging over sites for each month
```{r}

SoilBiomass_year <- dt1 %>%
  # group by year, treatment, and elevation, and soil horizon
  group_by(Year, Treatment, El, Hor) %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)),
            .groups = "drop") %>% 
  mutate(Elevation_num = case_when(
    El == "SF" ~ 790,
    El == "H" ~ 750,
    El == "U" ~ 675,
    El == "M" ~ 600,
    El == "L" ~ 525
  ))

SoilBiomass_year$Hor <- factor(SoilBiomass_year$Hor,
                               levels = c("Oi/Oe", "Oa/A", "Min"),
                               ordered = TRUE)

summary(SoilBiomass_year)
```

#Graph your analytes by month and year in ggplot
In this example, we have DIC and DOC
```{r}
SoilBioMass_yr <- ggplot(SoilBiomass_year, aes(x = Year, y = BIOC, 
                                               color = as.factor(Elevation_num))) +
  geom_point(size = 1) +
  facet_wrap(vars(Treatment, Hor)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Microbial biomass (mg/kg)",
    title = "Yearly microbial biomas over time colored by elevation, grouped \n
    by soil horizon and watershed (WW6 and W1)",
    color = "Elevation (m)") +
  theme_minimal()

SoilResp_yr <- ggplot(SoilBiomass_year, aes(x = Year, y = RESPC, 
                                   color = as.factor(Elevation_num))) +
  geom_point(size = 1) +
  facet_wrap(vars(Treatment, Hor)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Soil respiration rate (mg/kg/day)",
    title = "Yearly soil respiration rate over time colored by elevation, grouped \n
    by soil horizon and watershed (WW6 and W1)",
    color = "Elevation (m)") +
  theme_minimal()

SoilBioMassN_yr <- ggplot(SoilBiomass_year, aes(x = Year, y = BION, 
                                                color = as.factor(Elevation_num))) +
  geom_point(size = 1) +
  facet_wrap(vars(Treatment, Hor)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Microbial biomass N (mg/kg)",
    title = "Yearly microbial biomas N over time colored by elevation, grouped \n
    by soil horizon and watershed (WW6 and W1)",
    color = "Elevation (m)") +
  theme_minimal()

SoilNitrate_yr <- ggplot(SoilBiomass_year, aes(x = Year, y = NO3, 
                                               color = as.factor(Elevation_num))) +
  geom_point(size = 1) +
  facet_wrap(vars(Treatment, Hor)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Soil nitrate (mg/kg)",
    title = "Yearly soil nitrate over time colored by elevation, grouped \n
    by soil horizon and watershed (WW6 and W1)",
    color = "Elevation (m)") +
  theme_minimal()

SoilAmmonium_yr <- ggplot(SoilBiomass_year, aes(x = Year, y = NH4, 
                                               color = as.factor(Elevation_num))) +
  geom_point(size = 1) +
  facet_wrap(vars(Treatment, Hor)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Soil ammonium (mg/kg)",
    title = "Yearly soil ammonium over time colored by elevation, grouped \n
    by soil horizon and watershed (WW6 and W1)",
    color = "Elevation (m)") +
  theme_minimal()

SoilNIT_yr <- ggplot(SoilBiomass_year, aes(x = Year, y = NIT, 
                                               color = as.factor(Elevation_num))) +
  geom_point(size = 1) +
  facet_wrap(vars(Treatment, Hor)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Soil potential net nitrification (mg/kg/day)",
    title = "Yearly soil potential net nitrification over time colored by elevation, grouped \n
    by soil horizon and watershed (WW6 and W1)",
    color = "Elevation (m)") +
  theme_minimal()

SoilMIN_yr <- ggplot(SoilBiomass_year, aes(x = Year, y = MIN, 
                                               color = as.factor(Elevation_num))) +
  geom_point(size = 1) +
  facet_wrap(vars(Treatment, Hor)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Soil potential net N mineralization (mg/kg/day)",
    title = "Yearly soil potential net N mineralization over time colored by elevation, grouped \n
    by soil horizon and watershed (WW6 and W1)",
    color = "Elevation (m)") +
  theme_minimal()

SoilDEA_yr <- ggplot(SoilBiomass_year, aes(x = Year, y = DEA, 
                                               color = as.factor(Elevation_num))) +
  geom_point(size = 1) +
  facet_wrap(vars(Treatment, Hor)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Soil denitrification enzyme activity (unitless)",
    title = "Yearly soil denitrification enzyme activity over time colored by elevation, grouped \n
    by soil horizon and watershed (WW6 and W1)",
    color = "Elevation (m)") +
  theme_minimal()

SoilH2O_yr <- ggplot(SoilBiomass_year, aes(x = Year, y = H2O, 
                                               color = as.factor(Elevation_num))) +
  geom_point(size = 1) +
  facet_wrap(vars(Treatment, Hor)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Gravimetric water content (g/g)",
    title = "Yearly gravimetric water content over time colored by elevation, grouped \n
    by soil horizon and watershed (WW6 and W1)",
    color = "Elevation (m)") +
  theme_minimal()

SoilpH_yr <- ggplot(SoilBiomass_year, aes(x = Year, y = pH, 
                                               color = as.factor(Elevation_num))) +
  geom_point(size = 1) +
  facet_wrap(vars(Treatment, Hor)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Soil pH",
    title = "Yearly soil pH over time colored by elevation, grouped \n
    by soil horizon and watershed (WW6 and W1)",
    color = "Elevation (m)") +
  theme_minimal()

```

#Save all your plots to the graph folder
```{r}
plots <- list(SoilBioMass_yr, SoilResp_yr, SoilBioMassN_yr, SoilNitrate_yr,
              SoilAmmonium_yr, SoilNIT_yr, SoilMIN_yr, SoilDEA_yr, SoilH2O_yr,
              SoilpH_yr) 
names <- c("SoilBioMass_yr", "SoilResp_yr", "SoilBioMassN_yr", "SoilNitrate_yr",
           "SoilAmmonium_yr", "SoilNIT_yr", "SoilMIN_yr", "SoilDEA_yr", "SoilH2O_yr",
           "SoilpH_yr") 

for (i in seq_along(plots)) {
  ggsave(
    filename = here::here("Graphs", paste0(names[i], ".png")),
    plot = plots[[i]],
    width = 8,
    height = 5,
    dpi = 300
  )
}
```

LASTLY push your changes to our github repo!!