---
title: "HB Temporal C Synthesis: Aboveground Biomass"
author: "Caitlin Hicks Pries, Emerson Conrad-Rooney"
date: "`r Sys.Date()`"
output: html_document
---
#Setup
```{r setup, include=FALSE}
rm(list = ls()) #clear R environment

# Check for and install required packages and load libraries
for (package in c( 'tidyverse', 'here', 'pak', 'plotrix')) {    #can add more packages to this vector
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}

knitr::opts_knit$set(root.dir = here())
```

#Set up: HubbardBrookForestAnalytics package
https://github.com/kearutherford/HubbardBrookForestAnalytics
This package has internal data from Hubbard Brook
```{r install HubbardBrookForestAnalytics package}
# Install and load HubbardBrookForestAnalytics
pak::pak('kearutherford/HubbardBrookForestAnalytics')
library(HubbardBrookForestAnalytics)
```

#TITLES FROM EDI: 
#Hubbard Brook Experimental Forest: Watershed 1 Tree Inventory, 1996 - ongoing
DESCRIBE DATA: W6 aboveground biomass data from tree inventories every five years from 1996-2021.
https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-hbr.446.1

#Hubbard Brook Experimental Forest: Watershed 5 Tree Inventory, 1982 - ongoing
DESCRIBE DATA: W6 aboveground biomass data from tree inventories every five years from 1965-2022.
https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-hbr.447.1

#Hubbard Brook Experimental Forest: Watershed 6 Tree Inventory, 1965 - ongoing
DESCRIBE DATA: W6 aboveground biomass data from tree inventories every five years from 1965-2019.
https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-hbr.448.1

#Dealing with NA codes
~No NA values

#Access data from EDI
The data (EDI URLs listed above) are interally included in the HubbardBrookForestAnalytics package has these data internally, so this code does not have to access the data from EDI.

#Clean up your data
This is done by the HubbardBrookForestAnalytics package 

#Filter data to the window of time we are interested in
For now, keep it to its native time frame

#Summarize your data by year for graphing
DESCRIBE timestep of your data here: Yearly

```{r}
#Use HBEFBiomass() function to get internal data (data_type = "internal") summarized by plot (results = "by_plot")
HBEFBiomass_by_plot <- HBEFBiomass(data_type = "internal",
                               results = "by_plot")

#Summarize mean aboveground biomass by year and watershed
HBEFBiomass_summary <- HBEFBiomass_by_plot %>%
  group_by(year, watershed) %>%
  summarize(above_L_Mg_ha_mean = mean(above_L_Mg_ha),
            above_L_Mg_ha_SE = std.error(above_L_Mg_ha))

#Change Year to be numeric
HBEFBiomass_summary$year <- as.numeric(HBEFBiomass_summary$year)
```

#Graph your data by year (~5 years) in ggplot
```{r}
HBEFBiomass_W1 <- 
  ggplot(HBEFBiomass_summary[HBEFBiomass_summary$watershed == "W1",], aes(x = year, y = above_L_Mg_ha_mean)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Aboveground live tree biomass (Mg/ha)",
    title = "Watershed 1: Aboveground Living Biomass for Trees and Saplings") +
  geom_errorbar(aes(x = year, ymin=above_L_Mg_ha_mean-above_L_Mg_ha_SE, ymax=above_L_Mg_ha_mean+above_L_Mg_ha_SE), width=0.5) +
    ylim(0,225) +
  theme_minimal()

HBEFBiomass_W5 <- 
  ggplot(HBEFBiomass_summary[HBEFBiomass_summary$watershed == "W5",], aes(x = year, y = above_L_Mg_ha_mean)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Aboveground live tree biomass (Mg/ha)",
    title = "Watershed 5: Aboveground Living Biomass for Trees and Saplings") +
  geom_errorbar(aes(x = year, ymin=above_L_Mg_ha_mean-above_L_Mg_ha_SE, ymax=above_L_Mg_ha_mean+above_L_Mg_ha_SE), width=0.5) +
    ylim(0,210) +
  theme_minimal()

HBEFBiomass_W6 <- 
  ggplot(HBEFBiomass_summary[HBEFBiomass_summary$watershed == "W6",], aes(x = year, y = above_L_Mg_ha_mean)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Aboveground live tree biomass (Mg/ha)",
    title = "Watershed 6: Aboveground Living Biomass for Trees and Saplings") +
  geom_errorbar(aes(x = year, ymin=above_L_Mg_ha_mean-above_L_Mg_ha_SE, ymax=above_L_Mg_ha_mean+above_L_Mg_ha_SE), width=0.5) +
    ylim(0,200) +
  theme_minimal()

HBEFBiomass_W6 <- 
  ggplot(HBEFBiomass_summary[HBEFBiomass_summary$watershed == "W6",], aes(x = year, y = above_L_Mg_ha_mean)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Aboveground live tree biomass (Mg/ha)",
    title = "Watershed 6: Aboveground Living Biomass for Trees and Saplings") +
  geom_errorbar(aes(x = year, ymin=above_L_Mg_ha_mean-above_L_Mg_ha_SE, ymax=above_L_Mg_ha_mean+above_L_Mg_ha_SE), width=0.5) +
    ylim(0,200) +
  theme_minimal()

HBEFBiomass_W1.W5.W6 <- 
  ggplot(HBEFBiomass_summary, aes(x = year, y = above_L_Mg_ha_mean, color = watershed)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Aboveground live tree biomass (Mg/ha)",
    title = "Aboveground Living Biomass for Trees and Saplings") +
  geom_errorbar(aes(x = year, ymin=above_L_Mg_ha_mean-above_L_Mg_ha_SE, ymax=above_L_Mg_ha_mean+above_L_Mg_ha_SE), width=0.5) +
    ylim(0,225) +
  theme_minimal()
```

#Save all your plots to the graph folder
```{r}
plots <- list(HBEFBiomass_W1, HBEFBiomass_W5, HBEFBiomass_W6, HBEFBiomass_W1.W5.W6)
names <- c("Aboveground_Biomass_W1", "Aboveground_Biomass_W4", "Aboveground_Biomass_W6", "Aboveground_Biomass_W1.W5.W6") 

for (i in seq_along(plots)) {
  ggsave(
    filename = here::here("Graphs", paste0(names[i], ".png")),
    plot = plots[[i]],
    width = 8,
    height = 5,
    dpi = 300
  )
}
```

LASTLY push your changes to our github repo!!