---
title: "HB Temporal C Synthesis"
author: "Caitlin Hicks Pries"
date: "`r Sys.Date()`"
output: html_document
---
#Setup
```{r setup, include=FALSE}
rm(list = ls()) #clear R environment

# Check for and install required packages and load libraries
for (package in c( 'tidyverse', 'here')) {    #can add more packages to this vector
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}

knitr::opts_knit$set(root.dir = here())

```

#TITLE FROM EDI: Longitudinal Stream Chemistry at the Hubbard Brook Experimental Forest, Watershed 6, 1982 - ongoing
DESCRIBE DATA: W6 stream chemistry concentrations, longitudinal (not core weekly data)

#Dealing with NA codes
DESCRIBE the NA system for your dataset (below is example from Nina's work on the stream chemistry data):

Here are the descriptions of missing value codes from the metadata. The codes are the same for all variables, so that makes things easy. In this script, I code -999.99, -888.88 and -666.66 as NA when I load the data. Then I deal with -777.77 separately. 

-999.99, Data missing, or not taken at this time
-888.88, Sample volume too small for analysis
-666.66, Result Not Reported: Failed Internal QA/QC Review
-777.77, Analyte concentration below limit of detection

#Access data from EDI
Note where you need to paste in the url of your dataset
Note where you need to add in NA codes that you described above for your dataset

```{r access data from EDI}
options(HTTPUserAgent="EDI_CodeGen")

inUrl1  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/127/15/cfecb5b28e8f5d476b5d237e711f364a" #paste url here
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")

dt1 <-read.csv(infile1, stringsAsFactors=F, na.strings = c("-999.99", "-888.88", "-666.66")) #add in NA codes here!
unlink(infile1)

rm(inUrl1, infile1)
```

#Clean up your data
DESCRIBE how you clean up your data (below is example from Nina's work on the stream chemistry data).
YOU will have to convert the Date field properly using as.Date. If date is stored in a YYYY-MM-DD format, no need to specify the format. If not, scroll down to bottom of this page to see how to describe various formats: https://www.r-bloggers.com/2013/08/date-formats-in-r/

We will need to substitute half the detection limit for these data to avoid '-777.77' being treated as a numeric value. Note that if were to use the automatically generated EDI code, '-777.77' would become NA. Also, some values in the data table are below the stated detection limit for the analyte even though they do not have a -777.77 code, so with this step we are only dealing with the -777.77 but not making any decision about other numeric values that are below the limit. 

#Description of detection limits from the methods metadata: "Current detection limits for the Driscoll laboratory: 
Na (0.015 umol/L), K (0.12 umol/L), Ca (0.030 umol/L), 
Mg (0.0032 umol/L), Alt (0.0026 umol/L), Alm (1.3 umol/L), 
Alo (1.3 umol/L), NH4 (0.55 umol/L), H4SiO4 (3 umol/L), 
Cl (0.55 umol/L), NO3 (0.85 umol/L), SO4 (0.53 umol/L), 
F (0.097 umol/L), DIC (20 umol/L), DOC (20 umol/L), 
TN (0.81 umol/L). It is suggested to dataset users to use a value of one half the detection limit when a value for a solute is less than the detection limit.

#Make the changes.
As you will see below, in this example some judgement calls had to be made

```{r clean up data}

#Make a 'look-up' list of values to give measurements below detection limit
detectionLimits_Driscoll <- list(
  Na = 0.015/2,
  K = 0.12/2,
  Ca2 = 0.030/2,
  Mg2 = 0.0032/2,
  Alt = 0.0026/2,
  Altm = 1.3/2,
  Alom = 1.3/2,
  NH4 = 0.55/2,
  H4SiO4 = 3/2,
  NO3 = 0.85/2,
  SO4 = 0.53/2,
  F = 0.097/2,
  DIC = 20/2,
  DOC = 20/2,
  TN = 0.81/2,
  Cl = 0.55/2
)

W6_streamconc <- dt1 %>%
  mutate(Date = as.Date(Date)) %>% #convert to Date object
  mutate(across(  #replace -777.77 (values below the detection limit) with half the limit
    .cols = names(detectionLimits_Driscoll),
    .fns = ~ ifelse(.x == -777.77, detectionLimits_Driscoll[[cur_column()]], .x)))

#take a look to make sure the data is behaving how you want it to
summary(W6_streamconc) 

#Note that total monomeric Al (Altm) still has some values below -100. Deal with them:
unique(W6_streamconc$Altm[which(W6_streamconc$Altm < -100)]) 
#They are:
#[1] -1999.98  -887.01  -886.78  -886.64 -1777.76  -888.87
#take a closer look
#View(W6_longitudinal[which(W6_longitudinal$Altm < -100),])

#Change the Altm with values < -100 to NA and then remove rows from the df with no chemistry data for any analyte:
W6_streamconc <- W6_streamconc %>%
  mutate(Altm = ifelse(Altm < -100, NA, Altm)) %>%
  filter(!if_all(pH:TN, is.na))

#look at the data again
str(W6_streamconc)  
summary(W6_streamconc)  

#clean up
rm(dt1) #remove the original uncleaned data file

```

#Filter data to the window of time we are interested in
For now, keep it to its native time frame
# ```{r}
# W6_streamconc <- W6_streamconc %>% 
#   filter(Date > "1980-01-01" & Date< "2022-12-31")
# ```

#Summarize your data by month and year for graphing
DESCRIBE timestep(s) of your data here: Monthly
In the example below we are averaging over sites for each month
```{r}
W6_streamconc_month <- W6_streamconc %>%
  mutate(month = floor_date(Date, "month")) %>% #by month, converts all dates to 1st of month
  group_by(month) %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)),
            .groups = "drop")

W6_streamconc_year <- W6_streamconc %>%
  mutate(year = floor_date(Date, "year")) %>% #by year, converts all dates to 1st of year
  group_by(year) %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)),
            .groups = "drop")

#save these?? Right now no, and there is no data folder in the project
#write.csv(data_name, here("data", "data_name.csv"))
```

#Graph your analytes by month and year in ggplot
In this example, we have DIC and DOC
```{r}
DIC_month <- ggplot(W6_streamconc_month, aes(x = month, y = DIC)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Month",
    y = "Mean DIC (umol/L)",
    title = "Monthly Mean DIC Over Time") +
  theme_minimal()

DOC_month <- ggplot(W6_streamconc_month, aes(x = month, y = DOC)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Month",
    y = "Mean DOC (umol/L)",
    title = "Monthly Mean DOC Over Time") +
  theme_minimal()

DIC_year <- ggplot(W6_streamconc_year, aes(x = year, y = DIC)) +
  geom_point(size = 2) +
  geom_line() +
  labs(
    x = "Year",
    y = "Mean DIC (umol/L)",
    title = "Yearly Mean DIC Over Time") +
  theme_minimal()

DOC_year <- ggplot(W6_streamconc_year, aes(x = year, y = DOC)) +
  geom_point(size = 2) +
  geom_line() +
  labs(
    x = "Year",
    y = "Mean DOC (umol/L)",
    title = "Yearly Mean DOC Over Time") +
  theme_minimal()
```

#Save all your plots to the graph folder
```{r}
plots <- list(DIC_month, DOC_month, DIC_year, DOC_year) #change to your graph names
names <- c("DIC_month", "DOC_month", "DIC_year", "DOC_year") #change to names

for (i in seq_along(plots)) {
  ggsave(
    filename = here::here("Graphs", paste0(names[i], ".png")),
    plot = plots[[i]],
    width = 8,
    height = 5,
    dpi = 300
  )
}
```

LASTLY push your changes to our github repo!!