---
title: "HB Temporal C Synthesis"
author: "Caitlin Hicks Pries"
date: "`r Sys.Date()`"
output: html_document
---
#Setup
```{r setup, include=FALSE}
rm(list = ls()) #clear R environment

# Check for and install required packages and load libraries
for (package in c( 'tidyverse', 'here')) {    #can add more packages to this vector
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}

knitr::opts_knit$set(root.dir = here())

```

#DESCRIBE DATA HERE: Hubbard Brook Routine Seasonal Phenology Measurements 1989-present 

#Dealing with NA codes
DESCRIBE the NA system for your dataset 

There do not appear to be any NAs

#Access data from EDI
Note where you need to paste in the url of your dataset
Note where you need to add in NA codes that you described above for your dataset

```{r access data from EDI}
options(HTTPUserAgent="EDI_CodeGen")

inUrl1  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/51/16/9f623c83fb1da7595c6d2d498bde15df" 
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl"))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")

hb_phenology <-read.csv(infile1,header=F 
                   ,skip=1
                   ,sep=","  
                   , col.names=c(
                     "YEAR",     
                     "DATE",
                     "DAY",
                     "SEASON",
                     "SPECIES",
                     "SITE",
                     "Phenology State"), check.names=TRUE)

unlink(infile1)

```

#Data cleaning
Three species are observed across nine sites. 1B = bottom of W1; 4B = bottom of W4; 4T = top of W4; 5B = bottom of W5; 6T = top of W6; 7B = bottom of W7; 7T = top of W7; HQ = headquarters. Note that HQ was added in 1993, so there are 8 sites from 1989-1992

Observations are for phenological stages in the spring and in the fall that are enumerated as 1 -4 using a phenology index

Variables most relevant to a changing carbon cycle are likely timing of the start of the growing season, timing of of the end of the growing season, and growing season length

#For spring: 
1 = bud swelling, 2 = initial leaf expansion, 3 = leaves 1/2 final length, 4 = canopy in summer condition

#For fall: 
1  = no green in canopy, leaves 1/2 fallen, 2 = most leaves yellow or colored, most leaves on tree, 3 = many leaves still green, some yellow or coloring, 4 = canopy in summer condition

#Description of timing of spring greenup and fall green down: 
spring = first day that the phenology index >=3
fall = first day that the phenology index is <=3

This makes some assumptions about when photosynthesis turns and off

#Make the changes.
As you will see below, in this example some judgement calls had to be made

```{r access data from EDI}

#define the phenology index threshold that delineates the start and the end of the growing season
phenology_index_threshold <- 3

#create unique ID that includes YEAR, SPECIES, and SITE for identifying the first and last days that of exceeding the phenology threshold
hb_phenology$ID = paste(hb_phenology$YEAR, hb_phenology$SPECIES, hb_phenology$SITE)

#subset to only include spring or fall since growing season start and end dates
#have to be determined for each season separately
spring_phenology = hb_phenology[hb_phenology$SEASON == "SPRING", ]
fall_phenology = hb_phenology[hb_phenology$SEASON == "FALL", ]

#for each year, site, and species, select the first instance that the phenology index was >= 3 for spring and <= 3 for fall

greenup = spring_phenology %>%
  group_by(ID) %>%
  filter(Phenology.State >= phenology_index_threshold) %>%
  slice(1) %>%
  ungroup()

greendown = fall_phenology %>%
  group_by(ID) %>%
  filter(Phenology.State <= phenology_index_threshold) %>%
  slice(1) %>%
  ungroup()

#rename columns and drop some (not needed anymore)
greenup <- greenup %>%
  select(
    everything(),
    spring_date = DATE,
    spring_doy = DAY,
    spring_index = Phenology.State,
    -SEASON
  )

greendown <- greendown %>%
  select(
    everything(),
    fall_date = DATE,
    fall_doy = DAY,
    fall_index = Phenology.State,
    -YEAR,
    -SEASON,
    -SPECIES,
    -SITE
  )

#merge greenup and greendown to enable calculation of growing season length
growing_season <- greenup %>%
  left_join(
    greendown,
    by = "ID"
  )

#calculate growing season length
growing_season$gs_length = growing_season$fall_doy - growing_season$spring_doy

#convert dates to Date objects
growing_season$spring_date = as.Date(growing_season$spring_date)
growing_season$fall_date = as.Date(growing_season$fall_date)

#take a look to make sure the data is behaving how you want it to
summary(growing_season) 

#clean up
rm(hb_phenology, spring_phenology, fall_phenology) #remove the original uncleaned data files

```

#Summarize your data by month and year for graphing
Note that this step is skipped here because data are already in an annual timestep so code is commented out for now

Instead, created subsets for plotting to separate data by site, keeping data together by watershed even if split into top (T) or bottom (B)

DESCRIBE timestep(s) of your data here: annual
In the example below we are averaging over sites for each month
```{r}

site_1 <- growing_season %>% 
  filter(SITE == "1B")

site_4 <- growing_season %>% 
  filter(SITE == "4B" | SITE == "4T")

site_5 <- growing_season %>% 
  filter(SITE == "5B" | SITE == "5T")

site_6 <- growing_season %>% 
  filter(SITE == "6T")

site_7 <- growing_season %>% 
  filter(SITE == "7B" | SITE == "7T")

site_HQ <- growing_season %>% 
  filter(SITE == "HQ")


#W6_streamconc_month <- W6_streamconc %>%
#  mutate(month = floor_date(Date, "month")) %>% #by month, converts all dates to 1st of month
#  group_by(month) %>%
#  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)),
#            .groups = "drop")

#W6_streamconc_year <- W6_streamconc %>%
#  mutate(year = floor_date(Date, "year")) %>% #by year, converts all dates to 1st of year
#  group_by(year) %>%
#  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)),
#            .groups = "drop")

#save these?? Right now no, and there is no data folder in the project
#write.csv(data_name, here("data", "data_name.csv"))
```

#Graph your data by year in ggplot
In this version, plots are grouped by watershed, with points for species
```{r}

#growing season start
gs_start_ws1 <-ggplot(site_1, aes(x= YEAR, y = spring_doy, color = SPECIES)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Day of Year",
    title = "Green Up Date Over Time W1") +
  theme_minimal()

gs_start_ws4 <-ggplot(site_4, aes(x= YEAR, y = spring_doy, color = SPECIES, linetype = SITE)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Day of Year",
    title = "Green Up Date Over Time W4") +
  theme_minimal()

gs_start_ws5 <-ggplot(site_5, aes(x= YEAR, y = spring_doy, color = SPECIES, linetype = SITE)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Day of Year",
    title = "Green Up Date Over Time W5") +
  theme_minimal()

gs_start_ws6 <-ggplot(site_1, aes(x= YEAR, y = spring_doy, color = SPECIES)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Day of Year",
    title = "Green Up Date Over Time W6") +
  theme_minimal()

gs_start_ws7 <-ggplot(site_5, aes(x= YEAR, y = spring_doy, color = SPECIES, linetype = SITE)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Day of Year",
    title = "Green Up Date Over Time W7") +
  theme_minimal()

gs_start_HQ <-ggplot(site_1, aes(x= YEAR, y = spring_doy, color = SPECIES)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Day of Year",
    title = "Green Up Date Over Time HQ") +
  theme_minimal()

#growing season end
gs_end_ws1 <-ggplot(site_1, aes(x= YEAR, y = fall_doy, color = SPECIES)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Day of Year",
    title = "Green Down Date Over Time W1") +
  theme_minimal()

gs_end_ws4 <-ggplot(site_4, aes(x= YEAR, y = fall_doy, color = SPECIES, linetype = SITE)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Day of Year",
    title = "Green Down Date Over Time W4") +
  theme_minimal()

gs_end_ws5 <-ggplot(site_5, aes(x= YEAR, y = fall_doy, color = SPECIES, linetype = SITE)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Day of Year",
    title = "Green Down Date Over Time W5") +
  theme_minimal()

gs_end_ws6 <-ggplot(site_1, aes(x= YEAR, y = fall_doy, color = SPECIES)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Day of Year",
    title = "Green Down Date Over Time W6") +
  theme_minimal()

gs_end_ws7 <-ggplot(site_5, aes(x= YEAR, y = fall_doy, color = SPECIES, linetype = SITE)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Day of Year",
    title = "Green Down Date Over Time W7") +
  theme_minimal()

gs_end_HQ <-ggplot(site_1, aes(x= YEAR, y = fall_doy, color = SPECIES)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Day of Year",
    title = "Green Down Date Over Time HQ") +
  theme_minimal()

#growing season length
gs_length_ws1 <-ggplot(site_1, aes(x= YEAR, y = gs_length, color = SPECIES)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Growing Season Length Over Time W1") +
  theme_minimal()

gs_length_ws4 <-ggplot(site_4, aes(x= YEAR, y = gs_length, color = SPECIES, linetype = SITE)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Growing Season Length Over Time W4") +
  theme_minimal()

gs_length_ws5 <-ggplot(site_5, aes(x= YEAR, y = gs_length, color = SPECIES, linetype = SITE)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Growing Season Length Over Time W5") +
  theme_minimal()

gs_length_ws6 <-ggplot(site_1, aes(x= YEAR, y = gs_length, color = SPECIES)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Growing Season Length Over Time W6") +
  theme_minimal()

gs_length_ws7 <-ggplot(site_5, aes(x= YEAR, y = gs_length, color = SPECIES, linetype = SITE)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Growing Season Length Over Time W7") +
  theme_minimal()

gs_length_HQ <-ggplot(site_1, aes(x= YEAR, y = gs_length, color = SPECIES)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Growing Season Length Over Time HQ") +
  theme_minimal()

```

#Save all your plots to the graph folder
```{r}

#make list of plots
plots <- list(gs_start_ws1, gs_start_ws4, gs_start_ws5, gs_start_ws6, gs_start_ws7, gs_start_HQ, gs_end_ws1, gs_end_ws4, gs_end_ws5, gs_end_ws6, gs_end_ws7, gs_end_HQ,
gs_length_ws1, gs_length_ws4, gs_length_ws5, gs_length_ws6, gs_length_ws7, gs_length_HQ) 

#provide names
names <- c("gs_start_ws1", "gs_start_ws4", "gs_start_ws5", "gs_start_ws6", "gs_start_ws7", "gs_start_HQ", "gs_end_ws1", "gs_end_ws4", "gs_end_ws5", "gs_end_ws6", "gs_end_ws7", "gs_end_HQ", "gs_length_ws1", "gs_length_ws4", "gs_length_ws5", "gs_length_ws6", "gs_length_ws7", "gs_length_HQ") 

for (i in seq_along(plots)) {
  ggsave(
    filename = here::here("Graphs", paste0(names[i], ".png")),
    plot = plots[[i]],
    width = 8,
    height = 5,
    dpi = 300
  )
}
```

LASTLY push your changes to our github repo!!