---
title: "HB Temporal C Synthesis - Litterfall"
author: "Alix Contosta"
date: "`r Sys.Date()`"
output: html_document
---
#Setup
```{r setup, include=FALSE}
rm(list = ls()) #clear R environment

# Check for and install required packages and load libraries
for (package in c( 'tidyverse', 'here', 'plotrix')) {    #can add more packages to this vector
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}

knitr::opts_knit$set(root.dir = here())

```

#DESCRIBE DATA HERE: Soil-atmosphere fluxes of carbon dioxide, nitrous oxide and methane on Watershed 1 and Bear Brook, 2002-2024 

#Dealing with NA codes
DESCRIBE the NA system for your dataset 

Metadata on EDI notes that -9999 indicates "no data available." There are also instances of NAs showing up as -999.99, -9999.00, and -9999.000

#Access data from EDI
Note where you need to paste in the url of your dataset
Note where you need to add in NA codes that you described above for your dataset

```{r access data from EDI}
options(HTTPUserAgent="EDI_CodeGen")

inUrl1  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/116/14/f9ff9689063d019e7f7fde8ab4d9a5cc" 
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl"))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")
```


```{r access data from EDI}
```


```{r access data from EDI}
hb_gasflux <-read.csv(infile1,header=F 
                         ,skip=1
                         ,sep=","  
                         , col.names=c(
                           "Date",
                           "Site",
                           "Elevation",
                           "Chamber",
                           "CO2_rate",
                           "N2O_rate",
                           "CH4_rate"), 
                         na.strings = c("-999.99", "-9999",  "-9999.00", "-9999.000"), check.names=TRUE)



unlink(infile1)

```

#Data cleaning
Average monhtly and annual fluxes have to be calculated for each gas species and for Bear Brook and W1


#Make the changes.

```{r access data from EDI}
#convert Date to a Date object
hb_gasflux$Date = as.Date(hb_gasflux$Date, "%Y-%m-%d")

#separate out year for calculating annual averages
hb_gasflux$Year = strftime(hb_gasflux$Date, "%Y")

#separate out mounth for calculating monthly averages
hb_gasflux$Month = strftime(hb_gasflux$Date, "%m")

#add total litterfall mass for each unique YEAR x TAG combination
tag_total = hb_litterfall %>%
  group_by(YEAR, SITE, ELEV, TAG) %>%
  summarise(total_litterfall = sum(DRY_MASS, na.rm = TRUE))

#calculate average and standard error litterfall mass for each site x year (std error
#incorporates differences among elevations)
avg_litter = tag_total %>%
group_by(YEAR, SITE) %>%
  summarise(avg_litter = mean(total_litterfall, na.rm = TRUE),
            se_litter = std.error(total_litterfall, na.rm = TRUE))

#take a look to make sure the data is behaving how you want it to
summary(avg_litter) 

#clean up
rm(hb_litterfall, tag_total) #remove the original uncleaned data files

```

#Summarize your data by month and year for graphing
Note that this step is skipped here because data are already in an annual timestep so code is commented out for now


DESCRIBE timestep(s) of your data here: annual

```{r}

#W6_streamconc_month <- W6_streamconc %>%
#  mutate(month = floor_date(Date, "month")) %>% #by month, converts all dates to 1st of month
#  group_by(month) %>%
#  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)),
#            .groups = "drop")

#W6_streamconc_year <- W6_streamconc %>%
#  mutate(year = floor_date(Date, "year")) %>% #by year, converts all dates to 1st of year
#  group_by(year) %>%
#  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)),
#            .groups = "drop")

#save these?? Right now no, and there is no data folder in the project
#write.csv(data_name, here("data", "data_name.csv"))
```

#Graph your data by year in ggplot
In this version, there is a single plot, points represent SITE
```{r}

#litterfall
litter <-ggplot(avg_litter, aes(x= YEAR, y = avg_litter, color = SITE)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(x = YEAR, ymin = avg_litter - se_litter, ymax =
                      avg_litter + se_litter), width=0.5)+ 
  labs(
    x = "Year",
    y = "Litter Mass (g)",
    title = "Annual Litterfall") +
  theme_minimal()



```

#Save all your plots to the graph folder
```{r}

#make list of plots
plots <- list(litter) 

#provide names
names <- c("litterfall") 

for (i in seq_along(plots)) {
  ggsave(
    filename = here::here("Graphs", paste0(names[i], ".png")),
    plot = plots[[i]],
    width = 8,
    height = 5,
    dpi = 300
  )
}
```

LASTLY push your changes to our github repo!!