---
title: "HB Temporal Carbon - W6 Lysimeter Data"
author: "Angela Possinger"
date: "`r Sys.Date()`"
output: html_document
---
#Setup
```{r setup, include=FALSE}
rm(list = ls()) #clear R environment

# Check for and install required packages and load libraries
for (package in c( 'tidyverse', 'here')) {    #can add more packages to this vector
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}

knitr::opts_knit$set(root.dir = here())

```

#TITLE FROM EDI: Chemistry of freely-draining soil solutions at the Hubbard Brook Experimental Forest, Watershed 6, 1984 - present
DESCRIBE DATA: W6 lysimeter soil solution chemistry data, approximately monthly since 1984 
Focus variables: pH, ANC, Al, Si, NH4, NO3, DOC, DIC, TN

#Dealing with NA codes
Here are the descriptions of missing value codes from the metadata. The codes are the same for all variables, so that makes things easy. In this script, I code -999.99, -888.88 and -666.66 as NA when I load the data. Then I deal with -777.77 separately. 

-999.99, Data missing, or not taken at this time
-888.88, Sample volume too small for analysis
-666.66, Result Not Reported: Failed Internal QA/QC Review
-777.77, Analyte concentration below limit of detection

#Access data from EDI
Note where you need to paste in the url of your dataset
Note where you need to add in NA codes that you described above for your dataset

```{r access data from EDI}
options(HTTPUserAgent="EDI_CodeGen")

inUrl1  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/62/21/c03398425ac8e21a12750cc2655202e4" #paste url here
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")

dt1 <-read.csv(infile1, stringsAsFactors=F, na.strings = c("-999.99", "-888.88", "-666.66")) #add in NA codes here!
unlink(infile1)

rm(inUrl1, infile1)
```

#Clean up your data
DESCRIBE how you clean up your data (below is example from Nina's work on the stream chemistry data).
YOU will have to convert the Date field properly using as.Date. If date is stored in a YYYY-MM-DD format, no need to specify the format. If not, scroll down to bottom of this page to see how to describe various formats: https://www.r-bloggers.com/2013/08/date-formats-in-r/

We will need to substitute half the detection limit for these data to avoid '-777.77' being treated as a numeric value. Note that if were to use the automatically generated EDI code, '-777.77' would become NA. Also, some values in the data table are below the stated detection limit for the analyte even though they do not have a -777.77 code, so with this step we are only dealing with the -777.77 but not making any decision about other numeric values that are below the limit. 

Calculated variables such as ANC are not easily estimated by detection limits; -777.77 values for ANC will be coded as NA. 

pH doesn't have a logical detection limit; -777.77 values for pH will be coded as NA. 

#Description of detection limits from the methods metadata: "Current detection limits for the Driscoll laboratory: 
Na (0.015 umol/L), K (0.12 umol/L), Ca (0.030 umol/L), 
Mg (0.0032 umol/L), Alt (0.0026 umol/L), Alm (1.3 umol/L), 
Alo (1.3 umol/L), NH4 (0.55 umol/L), H4SiO4 (3 umol/L), 
Cl (0.55 umol/L), NO3 (0.85 umol/L), SO4 (0.53 umol/L), 
F (0.097 umol/L), DIC (20 umol/L), DOC (20 umol/L), 
TN (0.81 umol/L). It is suggested to dataset users to use a value of one half the detection limit when a value for a solute is less than the detection limit.

#Make the changes.
As you will see below, in this example some judgement calls had to be made

```{r clean up data}

#Remove "." from column names
names(dt1) <- gsub("\\.", "", names(dt1))

#Make a 'look-up' list of values to give measurements below detection limit
detectionLimits_Driscoll <- list(
  Na = 0.015/2,
  K = 0.12/2,
  Ca2 = 0.030/2,
  Mg2 = 0.0032/2,
  Alt = 0.0026/2,
  Altm = 1.3/2,
  Alom = 1.3/2,
  NH4 = 0.55/2,
  H4SiO4 = 3/2,
  NO3 = 0.85/2,
  SO4 = 0.53/2,
  F = 0.097/2,
  DIC = 20/2,
  DOC = 20/2,
  TN = 0.81/2,
  Cl = 0.55/2, 
  ANC = NA, 
  pH = NA
)

W6_lysimconc <- dt1 %>%
  mutate(Date = as.Date(Date)) %>% #convert to Date object
  mutate(across(  #replace -777.77 (values below the detection limit) with half the limit
    .cols = names(detectionLimits_Driscoll),
    .fns = ~ ifelse(.x == -777.77, detectionLimits_Driscoll[[cur_column()]], .x)))

#take a look to make sure the data is behaving how you want it to
summary(W6_lysimconc) 

#Extreme low values: pH, ANC, Al, NH4, NO3, DOC, DIC, TN
#pH
unique(W6_lysimconc$pH[which(W6_lysimconc$pH < 0)]) 
#They are:
#NA

#ANC
#Distribution of ANC 
hist(W6_lysimconc$ANC)
unique(W6_lysimconc$ANC[which(W6_lysimconc$ANC > 400)]) 
#They are:
#NA

#Alt
unique(W6_lysimconc$Alt[which(W6_lysimconc$Alt < -100)]) 
#NA

#Altm
unique(W6_lysimconc$Altm[which(W6_lysimconc$Altm < -100)]) 
#They are:
#-1999.98  -757.67 -1777.76  -773.38  -874.71
#take a closer look
View(W6_lysimconc[which(W6_lysimconc$Altm < -100),])
#All other values NA - set Altm to NA

#Alom
unique(W6_lysimconc$Alom[which(W6_lysimconc$Alom < -10)]) 
#They are:
#NA

#NH4
unique(W6_lysimconc$NH4[which(W6_lysimconc$NH4 < -100)]) 
#They are:
#NA

#NO3
unique(W6_lysimconc$NO3[which(W6_lysimconc$NO3 < -100)]) 
#They are:
#NA

#DOC
unique(W6_lysimconc$DOC[which(W6_lysimconc$DOC < -100)]) 
#They are:
#NA

#DIC
unique(W6_lysimconc$DIC[which(W6_lysimconc$DIC < -100)]) 
#They are:
#NA

#TN
unique(W6_lysimconc$TN[which(W6_lysimconc$TN < -10)]) 
#They are:
#NA

#Change the pH, Al, etc. with values defined above to NA and then remove rows from the df with no chemistry data for any analyte:
W6_lysimconc <- W6_lysimconc %>%
  #mutate(pH = ifelse(pH < 0, NA, pH)) %>%
  #mutate(Alt = ifelse(Alt < -100, NA, Alt)) %>%
  mutate(Altm = ifelse(Altm < -100, NA, Altm)) %>%
  #mutate(TN = ifelse(TN < -10, NA, TN)) %>%
  filter(!if_all(pH:TN, is.na))

#Convert the character columns to factors (needed for month and year function below)
W6_lysimconc <- W6_lysimconc %>%
  mutate(across(c(WS, Site, Horizon), as.factor))

#Add column for "organic" and "mineral" to split horizon types
W6_lysimconc <- W6_lysimconc %>%
  mutate(Horizon_type = ifelse(
    Horizon %in% c("Oe", "Oae", "Oa"),
    "Organic",
    "Mineral"
  ))


#clean up
rm(dt1) #remove the original uncleaned data file

```

#Filter data to the window of time we are interested in
For now, keep it to its native time frame
# ```{r}
# W6_streamconc <- W6_streamconc %>% 
#   filter(Date > "1980-01-01" & Date< "2022-12-31")
# ```

#Summarize your data by month and year for graphing
DESCRIBE timestep(s) of your data here: Monthly
In the example below we are averaging over sites for each month
```{r}
W6_lysimconc_month <- W6_lysimconc %>%
  mutate(month = floor_date(Date, "month")) %>% #by month, converts all dates to 1st of month
  group_by(month, WS, Horizon, Horizon_type) %>%
  summarise(
    across(
      where(is.numeric),
      list(
        mean = ~ mean(.x, na.rm = TRUE),
        se   = ~ sd(.x, na.rm = TRUE) / sqrt(sum(!is.na(.x)))
      ),
      .names = "{.col}_{.fn}"
    ),
    .groups = "drop"
  )

W6_lysimconc_year <- W6_lysimconc %>%
  mutate(year = floor_date(Date, "year")) %>% #by year, converts all dates to 1st of year
  group_by(year, WS, Horizon, Horizon_type) %>%
  summarise(
    across(
      where(is.numeric),
      list(
        mean = ~ mean(.x, na.rm = TRUE),
        se   = ~ sd(.x, na.rm = TRUE) / sqrt(sum(!is.na(.x)))
      ),
      .names = "{.col}_{.fn}"
    ),
    .groups = "drop"
  )

#save these?? Right now no, and there is no data folder in the project
#write.csv(data_name, here("data", "data_name.csv"))
```

#Graph your analytes by month and year in ggplot
```{r}
#Ordering for horizons
W6_lysimconc_month$Horizon <- factor(W6_lysimconc_month$Horizon, levels=c("Oa","Bh","Bs"))
W6_lysimconc_year$Horizon <- factor(W6_lysimconc_year$Horizon, levels=c("Oa","Bh","Bs"))

#pH#### 
#Month
pH_month <- ggplot(W6_lysimconc_month, aes(x = month, y = pH_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=pH_mean-pH_se, ymax=pH_mean+pH_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution pH",
    title = "W6 Monthly Mean Soil Solution pH Over Time") +
  theme_minimal()

#Year 
pH_year <- ggplot(W6_lysimconc_year, aes(x = year, y = pH_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=pH_mean-pH_se, ymax=pH_mean+pH_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution pH",
    title = "W6 Mean Soil Solution pH Over Time") +
  theme_minimal()

#ANC#### 
#Month
ANC_month <- ggplot(W6_lysimconc_month,
aes(x = month, y = ANC_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=ANC_mean-ANC_se, ymax=ANC_mean+ANC_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution ANC (microEquivalentPerLiter)",
    title = "W6 Monthly Mean Soil Solution ANC Over Time") +
  theme_minimal()

#Year 
ANC_year <- ggplot(W6_lysimconc_year,
aes(x = year, y = ANC_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=ANC_mean-ANC_se, ymax=ANC_mean+ANC_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution ANC (microEquivalentPerLiter)",
    title = "W6 Yearly Mean Soil Solution ANC Over Time") +
  theme_minimal()

#Alt#### 
#Month
Alt_month <- ggplot(W6_lysimconc_month, 
aes(x = month, y = Alt_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=Alt_mean-Alt_se, ymax=Alt_mean+Alt_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution Alt (micromolePerLiter)",
    title = "W6 Monthly Mean Soil Solution Total Al Over Time") +
  theme_minimal()

#Year 
Alt_year <- ggplot(W6_lysimconc_year, 
aes(x = year, y = Alt_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=Alt_mean-Alt_se, ymax=Alt_mean+Alt_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution Alt (micromolePerLiter)",
    title = "W6 Yearly Mean Soil Solution Total Al Over Time") +
  theme_minimal()

#Altm#### 
#Month
Altm_month <- ggplot(W6_lysimconc_month, 
aes(x = month, y = Altm_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=Altm_mean-Altm_se, ymax=Altm_mean+Altm_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution Alt (micromolePerLiter)",
    title = "W6 Monthly Mean Soil Solution Total Monomeric Al Over Time") +
  theme_minimal()

#Year 
Altm_year <- ggplot(W6_lysimconc_year,
aes(x = year, y = Altm_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=Altm_mean-Altm_se, ymax=Altm_mean+Altm_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution Altm (micromolePerLiter)",
    title = "W6 Yearly Mean Soil Solution Total Monomeric Al Over Time") +
  theme_minimal()

#Alom#### 
#Month
Alom_month <- ggplot(W6_lysimconc_month, 
aes(x = month, y = Alom_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=Alom_mean-Alom_se, ymax=Alom_mean+Alom_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution Alom (micromolePerLiter)",
    title = "W6 Monthly Mean Soil Solution Organic Monomeric Al Over Time") +
  theme_minimal()

#Year 
Alom_year <- ggplot(W6_lysimconc_year, 
aes(x = year, y = Alom_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=Alom_mean-Alom_se, ymax=Alom_mean+Alom_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution Alom (micromolePerLiter)",
    title = "W6 Yearly Mean Soil Solution Organic Monomeric Al Over Time") +
  theme_minimal()

#NH4#### 
#Month
NH4_month <- ggplot(W6_lysimconc_month,
aes(x = month, y = NH4_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=NH4_mean-NH4_se, ymax=NH4_mean+NH4_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution NH4+ (micromolePerLiter)",
    title = "W6 Monthly Mean Soil Solution NH4+ Over Time") +
  theme_minimal()

#Year 
NH4_year <- ggplot(W6_lysimconc_year, 
  aes(x = year, y = NH4_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=NH4_mean-NH4_se, ymax=NH4_mean+NH4_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution NH4+ (micromolePerLiter)",
    title = "W6 Yearly Mean Soil Solution NH4+ Over Time") +
  theme_minimal()

#NO3#### 
#Month
NO3_month <- ggplot(W6_lysimconc_month, 
  aes(x = month, y = NO3_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=NO3_mean-NO3_se, ymax=NO3_mean+NO3_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution NO3- (micromolePerLiter)",
    title = "W6 Monthly Mean Soil Solution NO3- Over Time") +
  theme_minimal()

#Year 
NO3_year <- ggplot(W6_lysimconc_year,
                           aes(x = year, y = NO3_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=NO3_mean-NO3_se, ymax=NO3_mean+NO3_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution NO3- (micromolePerLiter)",
    title = "W6 Yearly Mean Soil Solution NO3- Over Time") +
  theme_minimal()

#DOC#### 
#Month
DOC_month <- ggplot(W6_lysimconc_month, 
                    aes(x = month, y = DOC_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=DOC_mean-DOC_se, ymax=DOC_mean+DOC_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution DOC (micromolePerLiter)",
    title = "W6 Monthly Mean Soil Solution DOC Over Time") +
  theme_minimal()

#Year 
DOC_year <- ggplot(W6_lysimconc_year,
aes(x = year, y = DOC_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=DOC_mean-DOC_se, ymax=DOC_mean+DOC_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution DOC (micromolePerLiter)",
    title = "W6 Yearly Mean Soil Solution DOC Over Time") +
  theme_minimal()

#DIC#### 
#Month
DIC_month <- ggplot(W6_lysimconc_month, 
aes(x = month, y = DIC_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=DIC_mean-DIC_se, ymax=DIC_mean+DIC_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution DIC (micromolePerLiter)",
    title = "W6 Monthly Mean Soil Solution DIC Over Time") +
  theme_minimal()

#Year 
DIC_year <- ggplot(W6_lysimconc_year, 
aes(x = year, y = DIC_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=DIC_mean-DIC_se, ymax=DIC_mean+DIC_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution DIC (micromolePerLiter)",
    title = "W6 Yearly Mean Soil Solution DIC Over Time") +
  theme_minimal()

#TN#### 
#Month
TN_month <- ggplot(W6_lysimconc_month, 
aes(x = month, y = TN_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=TN_mean-TN_se, ymax=TN_mean+TN_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Month",
    y = "Soil solution TN (micromolePerLiter)",
    title = "W6 Monthly Mean Soil Solution TN Over Time") +
  theme_minimal()

#Year 
TN_year <- ggplot(W6_lysimconc_year,
aes(x = year, y = TN_mean)) +
  geom_point(size = 1) +
  geom_line() +
  geom_errorbar(aes(ymin=TN_mean-TN_se, ymax=TN_mean+TN_se), width=0.2) + 
  facet_grid(Horizon~.) + 
  labs(
    x = "Year",
    y = "Soil solution TN (micromolePerLiter)",
    title = "W6 Yearly Mean Soil Solution TN Over Time") +
  theme_minimal()

```

#Save all your plots to the graph folder
```{r}
plots <- list(
  pH_month, pH_year,
  ANC_month, ANC_year,
  Alt_month, Alt_year,
  Altm_month, Altm_year,
  Alom_month, Alom_year,
  NH4_month, NH4_year,
  NO3_month, NO3_year,
  DOC_month, DOC_year,
  DIC_month, DIC_year,
  TN_month, TN_year
)

#change to your graph names
names <- c(
  "W6_pH_month", "W6_pH_year",
  "W6_ANC_month", "W6_ANC_year",
  "W6_Alt_month","W6_Alt_year",
  "W6_Altm_month","W6_Altm_year",
  "W6_Alom_month","W6_Alom_year",
  "W6_NH4_month","W6_NH4_year",
  "W6_NO3_month","W6_NO3_year",
  "W6_DOC_month","W6_DOC_year",
  "W6_DIC_month","W6_DIC_year",
  "W6_TN_month","W6_TN_year"
)

for (i in seq_along(plots)) {
  ggsave(
    filename = here::here("Graphs", paste0(names[i], ".png")),
    plot = plots[[i]],
    width = 8,
    height = 10,
    dpi = 300
  )
}
```

LASTLY push your changes to our github repo!!