---
title: "HB Temporal C Synthesis: Aboveground Biomass for Bird Area"
author: "Emerson Conrad-Rooney, Caitlin Hicks Pries"
date: "`r Sys.Date()`"
output: html_document
---
#Setup
```{r setup, include=FALSE}
rm(list = ls()) #clear R environment

# Check for and install required packages and load libraries
for (package in c( 'tidyverse', 'here','plotrix')) {    #can add more packages to this vector
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}

knitr::opts_knit$set(root.dir = here())
```


#TITLES FROM EDI: 
#Ten-year interval tree-remeasure of a Northern Hardwood Forest: Bird Area, 1981 - ongoing, Hubbard Brook Experimental Forest
DESCRIBE DATA: DBH measurements every 10 years for 7443 trees in the Bird Area from 1981 - 2021.
https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-hbr.399.1

#Dealing with NA codes
~There are no NA values

#Access data from EDI
```{r access data from EDI}
options(HTTPUserAgent="EDI_CodeGen")
	      
inUrl1  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/399/1/220a0a3da8de0d36a6945b2f1d4d849c" 
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")

                   
Bird_area <-read.csv(infile1,header=F, skip=1, sep=",", quot='"',
               col.names=c("Plot", "Ln", "Species4", "DBH.cm", "Year", "PlotArea.ha", "isStatus"), check.names=TRUE)
               
unlink(infile1)
```

#Clean up your data
This is done by the HubbardBrookForestAnalytics package 
```{r}
#Calculate Basal Area from DBH
Bird_area$BasalArea.cm2 <- pi*((Bird_area$DBH.cm/2)^2)

#Summarize by plot for all living trees
Bird_area_summary <- Bird_area[Bird_area$isStatus != "Dead",] %>%
  group_by(Plot, Year, PlotArea.ha) %>%
  summarize(BasalArea.cm2_sum = sum(BasalArea.cm2))

#Convert Basal Area from cm2 to m2
Bird_area_summary$BasalArea.m2_sum <- Bird_area_summary$BasalArea.cm2_sum/10000

#Calculate BAm2/ha
Bird_area_summary$BasalArea.m2.ha <- Bird_area_summary$BasalArea.m2_sum/Bird_area_summary$PlotArea.ha
```

#Filter data to the window of time we are interested in
For now, keep it to its native time frame

#Summarize your data by year for graphing
DESCRIBE timestep of your data here: Yearly

```{r}
#Summarize mean BasalArea.m2.ha by year across plot
Bird_area_summary2 <- Bird_area_summary %>%
  group_by(Year) %>%
  summarize(BasalArea.m2.ha_mean = mean(BasalArea.m2.ha),
            BasalArea.m2.ha_SE = std.error(BasalArea.m2.ha))

#Change Year to be numeric
Bird_area_summary2$Year <- as.numeric(Bird_area_summary2$Year)
```

#Graph your data by year (~10 years) in ggplot
```{r}
HBEFBasalArea_BirdArea <- 
  ggplot(Bird_area_summary2, aes(x = Year, y = BasalArea.m2.ha_mean)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = "Basal Area (m2/ha)",
    title = "Bird Area: Basal Area for Living Trees") +
  geom_errorbar(aes(x = Year, ymin=BasalArea.m2.ha_mean-BasalArea.m2.ha_SE
, ymax=BasalArea.m2.ha_mean+BasalArea.m2.ha_SE
), width=0.5) +
  theme_minimal()

```

#Save all your plots to the graph folder
```{r}
plots <- list(HBEFBasalArea_BirdArea)
names <- c("Bird_Area_Basal_Area") 

for (i in seq_along(plots)) {
  ggsave(
    filename = here::here("Graphs", paste0(names[i], ".png")),
    plot = plots[[i]],
    width = 8,
    height = 5,
    dpi = 300
  )
}
```

LASTLY push your changes to our github repo!!