---
title: "HB Temporal C Synthesis: CCASE Aboveground Biomass Growth"
author: "Emerson Conrad-Rooney, Pamela Templer, Caitlin Hicks Pries"
date: "`r Sys.Date()`"
output: html_document
---

#Setup
```{r setup, include=FALSE}
rm(list = ls()) #clear R environment

# Check for and install required packages and load libraries
for (package in c( 'tidyverse', 'here','plotrix')) {    #can add more packages to this vector
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}

knitr::opts_knit$set(root.dir = here())
```


#TITLES FROM EDI: 
#Climate Change Across Seasons Experiment (CCASE) at the Hubbard Brook Experimental Forest: Tree Growth Data in support of "Declining Winter Snowpack Offsets Carbon Storage Enhancement from Growing Season Warming in Northern Temperate Forest Ecosystems", Conrad-Rooney et al. PNAS 2025DESCRIBE DATA: DBH measurements every 10 years for 7443 trees in the Bird Area from 1981 - 2021.
Annual stem biomass carbon growth for red maple trees at the Climate Change Across Seasons Experiment (CCASE)
https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-hbr.413.1

#Dealing with NA codes
~There are no NA values

#Access data from EDI
```{r access data from EDI}
options(HTTPUserAgent="EDI_CodeGen")
	      
inUrl1  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/413/1/6134ef924e83b940c5382d14dc30a1d1" 
infile2 <- tempfile()
try(download.file(inUrl1,infile1,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")

                   
CCASE_growth <-read.csv(infile1,header=F, skip=1, sep=",", quot='"',
               col.names=c("Plot", "Treatment", "Year", "Annual_C_growth_kgC.m2"), check.names=TRUE)
               
unlink(infile1)
```

#Filter data to the window of time we are interested in
For now, keep it to its native time frame

#Summarize your data by year for graphing
DESCRIBE timestep of your data here: Yearly

```{r}
#Summarize mean BasalArea.m2.ha by year and treatment across plot
CCASE_growth_summary <- CCASE_growth %>%
  group_by(Year, Treatment) %>%
  summarize(Annual_C_growth_kgC.m2_mean = mean(Annual_C_growth_kgC.m2),
            Annual_C_growth_kgC.m2_SE = std.error(Annual_C_growth_kgC.m2))
```

#Graph your data by year in ggplot
```{r}
CCASE_stem_growth <- 
  ggplot(CCASE_growth_summary, aes(x = Year, y = Annual_C_growth_kgC.m2_mean, color = Treatment)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Year",
    y = expression("Annual Stem Carbon Growth (kgC/m"^"2"*")"),
    title = "Climate Change Across Seasons Experiment (CCASE):\nRed Maple Aboveground Stem Carbon Growth") +
    scale_color_manual(values = c("black", "red3", "steelblue3")) +
  geom_errorbar(aes(x = Year, ymin=Annual_C_growth_kgC.m2_mean-Annual_C_growth_kgC.m2_SE, ymax=Annual_C_growth_kgC.m2_mean+Annual_C_growth_kgC.m2_SE
), width=0.2) +
  theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5))
```

#Save all your plots to the graph folder
```{r}
plots <- list(CCASE_stem_growth)
names <- c("CCASE_stem_growth") 

for (i in seq_along(plots)) {
  ggsave(
    filename = here::here("Graphs", paste0(names[i], ".png")),
    plot = plots[[i]],
    width = 8,
    height = 5,
    dpi = 300
  )
}
```

LASTLY push your changes to our github repo!!