---
title: "HB Temporal C Synthesis"
author: "Caitlin Hicks Pries"
date: "`r Sys.Date()`"
output: html_document
---
#Setup
```{r setup, include=FALSE}
rm(list = ls()) #clear R environment

# Check for and install required packages and load libraries
for (package in c( 'tidyverse', 'here')) {    #can add more packages to this vector
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}

knitr::opts_knit$set(root.dir = here())

```

#TITLE FROM EDI: Continuous precipitation and stream chemistry data, Hubbard Brook Ecosystem Study, 1963 – ongoing
DESCRIBE DATA: Precipitation chemistry, core HBWatER weekly sampling. (Data set also includes stream chem.)

#Dealing with NA codes
NAs are indicated as NA in the data, according to the metadata.

#Access data from EDI
Note where you need to paste in the url of your dataset
Note where you need to add in NA codes that you described above for your dataset

```{r access data from EDI}
options(HTTPUserAgent="EDI_CodeGen")

inUrl1  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/208/14/f148543b93982fd52f7089d2f2f51e85"
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")

dt1 <-read.csv(infile1,header=F 
          ,skip=1
            ,sep=","  
                ,quot='"' 
        , col.names=c(
                    "site",     
                    "date",     
                    "timeEST",     
                    "barcode",     
                    "pH",     
                    "DIC",     
                    "spCond",     
                    "temp",     
                    "ANC960",     
                    "ANCMet",     
                    "precipCatch",     
                    "fieldCode",     
                    "notes",     
                    "uniqueID",     
                    "waterYr",     
                    "Ca",     
                    "Mg",     
                    "K",     
                    "Na",     
                    "TMAl",     
                    "OMAl",     
                    "Al_ICP",     
                    "Al_ferron",     
                    "NH4",     
                    "SO4",     
                    "NO3",     
                    "Cl",     
                    "PO4",     
                    "DOC",     
                    "TDN",     
                    "DON",     
                    "SiO2",     
                    "Mn",     
                    "Fe",     
                    "F",     
                    "cationCharge",     
                    "anionCharge",     
                    "ionError",     
                    "duplicate",     
                    "sampleType",     
                    "ionBalance",     
                    "canonical",     
                    "pHmetrohm"    ), check.names=TRUE)
               
unlink(infile1)


rm(inUrl1, infile1)
```

#Clean up your data

# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
                
if (class(dt1$site)!="factor") dt1$site<- as.factor(dt1$site)                                   
# attempting to convert dt1$date dateTime string to R date structure (date or POSIXct)                                
tmpDateFormat<-"%Y-%m-%d"
tmp1date<-as.Date(dt1$date,format=tmpDateFormat)
# Keep the new dates only if they all converted correctly
if(nrow(dt1[dt1$date != "",]) == length(tmp1date[!is.na(tmp1date)])){dt1$date <- tmp1date } else {print("Date conversion failed for dt1$date. Please inspect the data and do the date conversion yourself.")}                                                                    
                                
if (class(dt1$barcode)!="factor") dt1$barcode<- as.factor(dt1$barcode)
if (class(dt1$pH)=="factor") dt1$pH <-as.numeric(levels(dt1$pH))[as.integer(dt1$pH) ]               
if (class(dt1$pH)=="character") dt1$pH <-as.numeric(dt1$pH)
if (class(dt1$DIC)=="factor") dt1$DIC <-as.numeric(levels(dt1$DIC))[as.integer(dt1$DIC) ]               
if (class(dt1$DIC)=="character") dt1$DIC <-as.numeric(dt1$DIC)
if (class(dt1$spCond)=="factor") dt1$spCond <-as.numeric(levels(dt1$spCond))[as.integer(dt1$spCond) ]               
if (class(dt1$spCond)=="character") dt1$spCond <-as.numeric(dt1$spCond)
if (class(dt1$temp)=="factor") dt1$temp <-as.numeric(levels(dt1$temp))[as.integer(dt1$temp) ]               
if (class(dt1$temp)=="character") dt1$temp <-as.numeric(dt1$temp)
if (class(dt1$ANC960)=="factor") dt1$ANC960 <-as.numeric(levels(dt1$ANC960))[as.integer(dt1$ANC960) ]               
if (class(dt1$ANC960)=="character") dt1$ANC960 <-as.numeric(dt1$ANC960)
if (class(dt1$ANCMet)=="factor") dt1$ANCMet <-as.numeric(levels(dt1$ANCMet))[as.integer(dt1$ANCMet) ]               
if (class(dt1$ANCMet)=="character") dt1$ANCMet <-as.numeric(dt1$ANCMet)
if (class(dt1$precipCatch)=="factor") dt1$precipCatch <-as.numeric(levels(dt1$precipCatch))[as.integer(dt1$precipCatch) ]               
if (class(dt1$precipCatch)=="character") dt1$precipCatch <-as.numeric(dt1$precipCatch)
if (class(dt1$fieldCode)!="factor") dt1$fieldCode<- as.factor(dt1$fieldCode)
if (class(dt1$notes)!="factor") dt1$notes<- as.factor(dt1$notes)
if (class(dt1$uniqueID)!="factor") dt1$uniqueID<- as.factor(dt1$uniqueID)
if (class(dt1$Ca)=="factor") dt1$Ca <-as.numeric(levels(dt1$Ca))[as.integer(dt1$Ca) ]               
if (class(dt1$Ca)=="character") dt1$Ca <-as.numeric(dt1$Ca)
if (class(dt1$Mg)=="factor") dt1$Mg <-as.numeric(levels(dt1$Mg))[as.integer(dt1$Mg) ]               
if (class(dt1$Mg)=="character") dt1$Mg <-as.numeric(dt1$Mg)
if (class(dt1$K)=="factor") dt1$K <-as.numeric(levels(dt1$K))[as.integer(dt1$K) ]               
if (class(dt1$K)=="character") dt1$K <-as.numeric(dt1$K)
if (class(dt1$Na)=="factor") dt1$Na <-as.numeric(levels(dt1$Na))[as.integer(dt1$Na) ]               
if (class(dt1$Na)=="character") dt1$Na <-as.numeric(dt1$Na)
if (class(dt1$TMAl)=="factor") dt1$TMAl <-as.numeric(levels(dt1$TMAl))[as.integer(dt1$TMAl) ]               
if (class(dt1$TMAl)=="character") dt1$TMAl <-as.numeric(dt1$TMAl)
if (class(dt1$OMAl)=="factor") dt1$OMAl <-as.numeric(levels(dt1$OMAl))[as.integer(dt1$OMAl) ]               
if (class(dt1$OMAl)=="character") dt1$OMAl <-as.numeric(dt1$OMAl)
if (class(dt1$Al_ICP)=="factor") dt1$Al_ICP <-as.numeric(levels(dt1$Al_ICP))[as.integer(dt1$Al_ICP) ]               
if (class(dt1$Al_ICP)=="character") dt1$Al_ICP <-as.numeric(dt1$Al_ICP)
if (class(dt1$Al_ferron)=="factor") dt1$Al_ferron <-as.numeric(levels(dt1$Al_ferron))[as.integer(dt1$Al_ferron) ]               
if (class(dt1$Al_ferron)=="character") dt1$Al_ferron <-as.numeric(dt1$Al_ferron)
if (class(dt1$NH4)=="factor") dt1$NH4 <-as.numeric(levels(dt1$NH4))[as.integer(dt1$NH4) ]               
if (class(dt1$NH4)=="character") dt1$NH4 <-as.numeric(dt1$NH4)
if (class(dt1$SO4)=="factor") dt1$SO4 <-as.numeric(levels(dt1$SO4))[as.integer(dt1$SO4) ]               
if (class(dt1$SO4)=="character") dt1$SO4 <-as.numeric(dt1$SO4)
if (class(dt1$NO3)=="factor") dt1$NO3 <-as.numeric(levels(dt1$NO3))[as.integer(dt1$NO3) ]               
if (class(dt1$NO3)=="character") dt1$NO3 <-as.numeric(dt1$NO3)
if (class(dt1$Cl)=="factor") dt1$Cl <-as.numeric(levels(dt1$Cl))[as.integer(dt1$Cl) ]               
if (class(dt1$Cl)=="character") dt1$Cl <-as.numeric(dt1$Cl)
if (class(dt1$PO4)=="factor") dt1$PO4 <-as.numeric(levels(dt1$PO4))[as.integer(dt1$PO4) ]               
if (class(dt1$PO4)=="character") dt1$PO4 <-as.numeric(dt1$PO4)
if (class(dt1$DOC)=="factor") dt1$DOC <-as.numeric(levels(dt1$DOC))[as.integer(dt1$DOC) ]               
if (class(dt1$DOC)=="character") dt1$DOC <-as.numeric(dt1$DOC)
if (class(dt1$TDN)=="factor") dt1$TDN <-as.numeric(levels(dt1$TDN))[as.integer(dt1$TDN) ]               
if (class(dt1$TDN)=="character") dt1$TDN <-as.numeric(dt1$TDN)
if (class(dt1$DON)=="factor") dt1$DON <-as.numeric(levels(dt1$DON))[as.integer(dt1$DON) ]               
if (class(dt1$DON)=="character") dt1$DON <-as.numeric(dt1$DON)
if (class(dt1$SiO2)=="factor") dt1$SiO2 <-as.numeric(levels(dt1$SiO2))[as.integer(dt1$SiO2) ]               
if (class(dt1$SiO2)=="character") dt1$SiO2 <-as.numeric(dt1$SiO2)
if (class(dt1$Mn)=="factor") dt1$Mn <-as.numeric(levels(dt1$Mn))[as.integer(dt1$Mn) ]               
if (class(dt1$Mn)=="character") dt1$Mn <-as.numeric(dt1$Mn)
if (class(dt1$Fe)=="factor") dt1$Fe <-as.numeric(levels(dt1$Fe))[as.integer(dt1$Fe) ]               
if (class(dt1$Fe)=="character") dt1$Fe <-as.numeric(dt1$Fe)
if (class(dt1$F)=="factor") dt1$F <-as.numeric(levels(dt1$F))[as.integer(dt1$F) ]               
if (class(dt1$F)=="character") dt1$F <-as.numeric(dt1$F)
if (class(dt1$cationCharge)=="factor") dt1$cationCharge <-as.numeric(levels(dt1$cationCharge))[as.integer(dt1$cationCharge) ]               
if (class(dt1$cationCharge)=="character") dt1$cationCharge <-as.numeric(dt1$cationCharge)
if (class(dt1$anionCharge)=="factor") dt1$anionCharge <-as.numeric(levels(dt1$anionCharge))[as.integer(dt1$anionCharge) ]               
if (class(dt1$anionCharge)=="character") dt1$anionCharge <-as.numeric(dt1$anionCharge)
if (class(dt1$ionError)=="factor") dt1$ionError <-as.numeric(levels(dt1$ionError))[as.integer(dt1$ionError) ]               
if (class(dt1$ionError)=="character") dt1$ionError <-as.numeric(dt1$ionError)
if (class(dt1$duplicate)!="factor") dt1$duplicate<- as.factor(dt1$duplicate)
if (class(dt1$sampleType)!="factor") dt1$sampleType<- as.factor(dt1$sampleType)
if (class(dt1$ionBalance)=="factor") dt1$ionBalance <-as.numeric(levels(dt1$ionBalance))[as.integer(dt1$ionBalance) ]               
if (class(dt1$ionBalance)=="character") dt1$ionBalance <-as.numeric(dt1$ionBalance)
if (class(dt1$canonical)!="factor") dt1$canonical<- as.factor(dt1$canonical)
if (class(dt1$pHmetrohm)=="factor") dt1$pHmetrohm <-as.numeric(levels(dt1$pHmetrohm))[as.integer(dt1$pHmetrohm) ]               
if (class(dt1$pHmetrohm)=="character") dt1$pHmetrohm <-as.numeric(dt1$pHmetrohm)
                
# Convert Missing Values to NA for non-dates
                
dt1$site <- as.factor(ifelse((trimws(as.character(dt1$site))==trimws("NA")),NA,as.character(dt1$site)))
dt1$barcode <- as.factor(ifelse((trimws(as.character(dt1$barcode))==trimws("NA")),NA,as.character(dt1$barcode)))
dt1$pH <- ifelse((trimws(as.character(dt1$pH))==trimws("NA")),NA,dt1$pH)               
suppressWarnings(dt1$pH <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$pH))==as.character(as.numeric("NA"))),NA,dt1$pH))
dt1$DIC <- ifelse((trimws(as.character(dt1$DIC))==trimws("NA")),NA,dt1$DIC)               
suppressWarnings(dt1$DIC <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$DIC))==as.character(as.numeric("NA"))),NA,dt1$DIC))
dt1$spCond <- ifelse((trimws(as.character(dt1$spCond))==trimws("NA")),NA,dt1$spCond)               
suppressWarnings(dt1$spCond <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$spCond))==as.character(as.numeric("NA"))),NA,dt1$spCond))
dt1$temp <- ifelse((trimws(as.character(dt1$temp))==trimws("NA")),NA,dt1$temp)               
suppressWarnings(dt1$temp <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$temp))==as.character(as.numeric("NA"))),NA,dt1$temp))
dt1$ANC960 <- ifelse((trimws(as.character(dt1$ANC960))==trimws("NA")),NA,dt1$ANC960)               
suppressWarnings(dt1$ANC960 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$ANC960))==as.character(as.numeric("NA"))),NA,dt1$ANC960))
dt1$ANCMet <- ifelse((trimws(as.character(dt1$ANCMet))==trimws("NA")),NA,dt1$ANCMet)               
suppressWarnings(dt1$ANCMet <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$ANCMet))==as.character(as.numeric("NA"))),NA,dt1$ANCMet))
dt1$precipCatch <- ifelse((trimws(as.character(dt1$precipCatch))==trimws("NA")),NA,dt1$precipCatch)               
suppressWarnings(dt1$precipCatch <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$precipCatch))==as.character(as.numeric("NA"))),NA,dt1$precipCatch))
dt1$fieldCode <- as.factor(ifelse((trimws(as.character(dt1$fieldCode))==trimws("NA")),NA,as.character(dt1$fieldCode)))
dt1$notes <- as.factor(ifelse((trimws(as.character(dt1$notes))==trimws("NA")),NA,as.character(dt1$notes)))
dt1$uniqueID <- as.factor(ifelse((trimws(as.character(dt1$uniqueID))==trimws("NA")),NA,as.character(dt1$uniqueID)))
dt1$Ca <- ifelse((trimws(as.character(dt1$Ca))==trimws("NA")),NA,dt1$Ca)               
suppressWarnings(dt1$Ca <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$Ca))==as.character(as.numeric("NA"))),NA,dt1$Ca))
dt1$Mg <- ifelse((trimws(as.character(dt1$Mg))==trimws("NA")),NA,dt1$Mg)               
suppressWarnings(dt1$Mg <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$Mg))==as.character(as.numeric("NA"))),NA,dt1$Mg))
dt1$K <- ifelse((trimws(as.character(dt1$K))==trimws("NA")),NA,dt1$K)               
suppressWarnings(dt1$K <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$K))==as.character(as.numeric("NA"))),NA,dt1$K))
dt1$Na <- ifelse((trimws(as.character(dt1$Na))==trimws("NA")),NA,dt1$Na)               
suppressWarnings(dt1$Na <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$Na))==as.character(as.numeric("NA"))),NA,dt1$Na))
dt1$TMAl <- ifelse((trimws(as.character(dt1$TMAl))==trimws("NA")),NA,dt1$TMAl)               
suppressWarnings(dt1$TMAl <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$TMAl))==as.character(as.numeric("NA"))),NA,dt1$TMAl))
dt1$OMAl <- ifelse((trimws(as.character(dt1$OMAl))==trimws("NA")),NA,dt1$OMAl)               
suppressWarnings(dt1$OMAl <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$OMAl))==as.character(as.numeric("NA"))),NA,dt1$OMAl))
dt1$Al_ICP <- ifelse((trimws(as.character(dt1$Al_ICP))==trimws("NA")),NA,dt1$Al_ICP)               
suppressWarnings(dt1$Al_ICP <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$Al_ICP))==as.character(as.numeric("NA"))),NA,dt1$Al_ICP))
dt1$Al_ferron <- ifelse((trimws(as.character(dt1$Al_ferron))==trimws("NA")),NA,dt1$Al_ferron)               
suppressWarnings(dt1$Al_ferron <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$Al_ferron))==as.character(as.numeric("NA"))),NA,dt1$Al_ferron))
dt1$NH4 <- ifelse((trimws(as.character(dt1$NH4))==trimws("NA")),NA,dt1$NH4)               
suppressWarnings(dt1$NH4 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$NH4))==as.character(as.numeric("NA"))),NA,dt1$NH4))
dt1$SO4 <- ifelse((trimws(as.character(dt1$SO4))==trimws("NA")),NA,dt1$SO4)               
suppressWarnings(dt1$SO4 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$SO4))==as.character(as.numeric("NA"))),NA,dt1$SO4))
dt1$NO3 <- ifelse((trimws(as.character(dt1$NO3))==trimws("NA")),NA,dt1$NO3)               
suppressWarnings(dt1$NO3 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$NO3))==as.character(as.numeric("NA"))),NA,dt1$NO3))
dt1$Cl <- ifelse((trimws(as.character(dt1$Cl))==trimws("NA")),NA,dt1$Cl)               
suppressWarnings(dt1$Cl <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$Cl))==as.character(as.numeric("NA"))),NA,dt1$Cl))
dt1$PO4 <- ifelse((trimws(as.character(dt1$PO4))==trimws("NA")),NA,dt1$PO4)               
suppressWarnings(dt1$PO4 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$PO4))==as.character(as.numeric("NA"))),NA,dt1$PO4))
dt1$DOC <- ifelse((trimws(as.character(dt1$DOC))==trimws("NA")),NA,dt1$DOC)               
suppressWarnings(dt1$DOC <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$DOC))==as.character(as.numeric("NA"))),NA,dt1$DOC))
dt1$TDN <- ifelse((trimws(as.character(dt1$TDN))==trimws("NA")),NA,dt1$TDN)               
suppressWarnings(dt1$TDN <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$TDN))==as.character(as.numeric("NA"))),NA,dt1$TDN))
dt1$DON <- ifelse((trimws(as.character(dt1$DON))==trimws("NA")),NA,dt1$DON)               
suppressWarnings(dt1$DON <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$DON))==as.character(as.numeric("NA"))),NA,dt1$DON))
dt1$SiO2 <- ifelse((trimws(as.character(dt1$SiO2))==trimws("NA")),NA,dt1$SiO2)               
suppressWarnings(dt1$SiO2 <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$SiO2))==as.character(as.numeric("NA"))),NA,dt1$SiO2))
dt1$Mn <- ifelse((trimws(as.character(dt1$Mn))==trimws("NA")),NA,dt1$Mn)               
suppressWarnings(dt1$Mn <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$Mn))==as.character(as.numeric("NA"))),NA,dt1$Mn))
dt1$Fe <- ifelse((trimws(as.character(dt1$Fe))==trimws("NA")),NA,dt1$Fe)               
suppressWarnings(dt1$Fe <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$Fe))==as.character(as.numeric("NA"))),NA,dt1$Fe))
dt1$F <- ifelse((trimws(as.character(dt1$F))==trimws("NA")),NA,dt1$F)               
suppressWarnings(dt1$F <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$F))==as.character(as.numeric("NA"))),NA,dt1$F))
dt1$cationCharge <- ifelse((trimws(as.character(dt1$cationCharge))==trimws("NA")),NA,dt1$cationCharge)               
suppressWarnings(dt1$cationCharge <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$cationCharge))==as.character(as.numeric("NA"))),NA,dt1$cationCharge))
dt1$anionCharge <- ifelse((trimws(as.character(dt1$anionCharge))==trimws("NA")),NA,dt1$anionCharge)               
suppressWarnings(dt1$anionCharge <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$anionCharge))==as.character(as.numeric("NA"))),NA,dt1$anionCharge))
dt1$ionError <- ifelse((trimws(as.character(dt1$ionError))==trimws("NA")),NA,dt1$ionError)               
suppressWarnings(dt1$ionError <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$ionError))==as.character(as.numeric("NA"))),NA,dt1$ionError))
dt1$duplicate <- as.factor(ifelse((trimws(as.character(dt1$duplicate))==trimws("NA")),NA,as.character(dt1$duplicate)))
dt1$sampleType <- as.factor(ifelse((trimws(as.character(dt1$sampleType))==trimws("NA")),NA,as.character(dt1$sampleType)))
dt1$ionBalance <- ifelse((trimws(as.character(dt1$ionBalance))==trimws("NA")),NA,dt1$ionBalance)               
suppressWarnings(dt1$ionBalance <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$ionBalance))==as.character(as.numeric("NA"))),NA,dt1$ionBalance))
dt1$canonical <- as.factor(ifelse((trimws(as.character(dt1$canonical))==trimws("NA")),NA,as.character(dt1$canonical)))
dt1$pHmetrohm <- ifelse((trimws(as.character(dt1$pHmetrohm))==trimws("NA")),NA,dt1$pHmetrohm)               
suppressWarnings(dt1$pHmetrohm <- ifelse(!is.na(as.numeric("NA")) & (trimws(as.character(dt1$pHmetrohm))==as.character(as.numeric("NA"))),NA,dt1$pHmetrohm))

# Here is the structure of the input data frame:
print("dt1) Structure")		    
str(dt1)                            

```

#Summarize your data by month and year for graphing
DESCRIBE timestep(s) of your data here: Weekly, usually
In the example below we are averaging over sites for each month
```{r}

#Quick plot of data for all rain gauges
ggplot(dt1) +
geom_point(mapping=aes(x=date,y=SO4)) +
facet_wrap(vars(site))

#What intervals between measurements at RG22 are present in the data, and how
#many observations at each interval?
unique(diff(sort(dt1$date[dt1$site=='RG22'])))
table(diff(sort(dt1$date[dt1$site=='RG22'])))

```

#Graph your analytes by month and year in ggplot
In this example, we have DIC and DOC
```{r}
DIC_month <- ggplot(W6_streamconc_month, aes(x = month, y = DIC)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Month",
    y = "Mean DIC (umol/L)",
    title = "Monthly Mean DIC Over Time") +
  theme_minimal()

DOC_month <- ggplot(W6_streamconc_month, aes(x = month, y = DOC)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Month",
    y = "Mean DOC (umol/L)",
    title = "Monthly Mean DOC Over Time") +
  theme_minimal()

DIC_year <- ggplot(W6_streamconc_year, aes(x = year, y = DIC)) +
  geom_point(size = 2) +
  geom_line() +
  labs(
    x = "Year",
    y = "Mean DIC (umol/L)",
    title = "Yearly Mean DIC Over Time") +
  theme_minimal()

DOC_year <- ggplot(W6_streamconc_year, aes(x = year, y = DOC)) +
  geom_point(size = 2) +
  geom_line() +
  labs(
    x = "Year",
    y = "Mean DOC (umol/L)",
    title = "Yearly Mean DOC Over Time") +
  theme_minimal()
```

#Save all your plots to the graph folder
```{r}
plots <- list(DIC_month, DOC_month, DIC_year, DOC_year) #change to your graph names
names <- c("DIC_month", "DOC_month", "DIC_year", "DOC_year") #change to names

for (i in seq_along(plots)) {
  ggsave(
    filename = here::here("Graphs", paste0(names[i], ".png")),
    plot = plots[[i]],
    width = 8,
    height = 5,
    dpi = 300
  )
}
```

LASTLY push your changes to our github repo!!