---
title: "HB Temporal C Synthesis: Forest Floor W6"
author: "Caitlin Hicks Pries"
date: "`r Sys.Date()`"
output: html_document
---
#Setup
```{r setup, include=FALSE}
rm(list = ls()) #clear R environment

# Check for and install required packages and load libraries
for (package in c( 'tidyverse', 'here', 'plotrix')) {    #can add more packages to this vector
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}

knitr::opts_knit$set(root.dir = here())

```

#TITLE FROM EDI: Mass and Chemistry of Organic Horizons and Surface Mineral Soils on Watershed 6 at the Hubbard Brook Experimental Forest, 1976 - present.
DESCRIBE DATA: W6 forest floor chemistry and mass

#Dealing with NA codes
DESCRIBE the NA system for your dataset (below is example from Nina's work on the stream chemistry data):

Here are the descriptions of missing value codes from the metadata. The codes are the same for all variables, so that makes things easy. In this script, I code -999.99, -888.88 and -666.66 as NA when I load the data. Then I deal with -777.77 separately. 

-9999.9, -9999, Data missing, or not taken at this time


#Access data from EDI
Note where you need to paste in the url of your dataset
Note where you need to add in NA codes that you described above for your dataset

Bring in Data Files (there are 6!)
```{r access data from EDI}
options(HTTPUserAgent="EDI_CodeGen")

inUrl1  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/172/4/16f1fab6db79c79440fbb4c3ff2e7ed7" #paste url here
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")

dtSiteInfo <-read.csv(infile1, stringsAsFactors=F, na.strings = c("-9999.9", "-9999")) #add in NA codes here!
unlink(infile1)

inUrl2  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/172/4/50d1e45dc7c0f5769e42f6d972e86f12" 
infile2 <- tempfile()
try(download.file(inUrl2,infile2,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile2))) download.file(inUrl2,infile2,method="auto")

dtPlotLoc <-read.csv(infile2, stringsAsFactors=F, na.strings = c("-9999.9", "-9999")) #add in NA codes here!
unlink(infile2)

inUrl3  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/172/4/382acf1fdcb69da1cca6674929f5ad6b" 
infile3 <- tempfile()
try(download.file(inUrl3,infile3,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile3))) download.file(inUrl3,infile3,method="auto")

dtMetalperm2 <-read.csv(infile3, stringsAsFactors=F, na.strings = c("-9999.99", "-8888.88")) #add in NA codes here!
unlink(infile3)

inUrl4  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/172/4/0ff5d3e2ca711915795b03770f6e9608" 
infile4 <- tempfile()
try(download.file(inUrl4,infile4,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile4))) download.file(inUrl4,infile4,method="auto")

dtMetalperkg <-read.csv(infile4, stringsAsFactors=F, na.strings = c("-9999.9", "-8888.8")) #add in NA codes here!
unlink(infile4)

inUrl5  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/172/4/775e243b7a2b67c0047498533bf5b9d1" 
infile5 <- tempfile()
try(download.file(inUrl5,infile5,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile5))) download.file(inUrl5,infile5,method="auto")

dtMass <-read.csv(infile5, stringsAsFactors=F, na.strings = c("-9999.9","-9999.99","-8888.8")) #add in NA codes here!
unlink(infile5)

inUrl6  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/172/4/f25fc11474e2b787cecc67949ecd0028" 
infile6 <- tempfile()
try(download.file(inUrl6,infile6,method="curl",extra=paste0(' -A "',getOption("HTTPUserAgent"),'"')))
if (is.na(file.size(infile6))) download.file(inUrl6,infile6,method="auto")

dtCN <-read.csv(infile6, stringsAsFactors=F, na.strings = c("-9999.9", "-8888.8", "-99.99")) #add in NA codes here!
unlink(infile6)

rm(inUrl1, infile1, inUrl2, infile2, inUrl3, infile3, inUrl4, infile4, inUrl5, infile5)

str(dtSiteInfo)
str(dtPlotLoc)
```

#Clean up your data
DESCRIBE how you clean up your data (below is example from Nina's work on the stream chemistry data).
YOU will have to convert the Date field properly using as.Date. If date is stored in a YYYY-MM-DD format, no need to specify the format. If not, scroll down to bottom of this page to see how to describe various formats: https://www.r-bloggers.com/2013/08/date-formats-in-r/

There do not seem to be any issues with detection limits. The data clean up here is all about combining these dataframes and calculating useful data like the amount of forest floor OM, C, and N

#Make the changes.

```{r clean up data}
comb1 <- left_join(dtMass, dtCN, by = c( "Year", "Plot", "Horizon", "Watershed")) %>% 
  select(-Site_ID) %>% 
  mutate(OM_pool = OM_TM*PerCentC/100) #TM is in kg/m2 and PerCentC is in % so OM_pool is kg C/m2

  
dtSiteInfoLong <- dtSiteInfo %>% 
  rowwise() %>%
  mutate ( "FF" = mean(c(FF_Thickness1,FF_Thickness2,FF_Thickness3,FF_Thickness4,FF_Thickness5,FF_Thickness6,
                            FF_Thickness7, FF_Thickness8), na.rm=TRUE), #FF_Thick
          Oie = mean(c(Oie_Thickness1,Oie_Thickness2,Oie_Thickness3,Oie_Thickness4,Oie_Thickness5,
                             Oie_Thickness6,Oie_Thickness7, Oie_Thickness8),na.rm=TRUE),
          min = mean(c(Core_1, Core_2, Core_3, Core_4),na.rm=TRUE),
          Oa = FF- Oie) %>% 
  select(c("Year", "Plot","FF", "Oie", "min", "Oa" )) %>% 
  pivot_longer(cols = FF:Oa,
               names_to = "Horizon",
               values_to = "thick") %>% 
  mutate(Horizon = as.factor(Horizon))  

comb2 <- comb1 %>% 
  mutate(Horizon = case_when(Horizon == "Oie+a" ~ "FF", TRUE ~ Horizon),
         Horizon = as.factor(Horizon)) %>% 
  left_join(., dtSiteInfoLong, by = c("Year", "Plot", "Horizon"), relationship = "many-to-many")

elev_loc <- read.csv("Grid_Elevation_WS6.csv")

finalFF <- comb2 %>% 
  left_join(., elev_loc, by = "Plot")

```

#Filter data to the window of time we are interested in
For now, keep it to its native time frame
# ```{r}
# W6_streamconc <- W6_streamconc %>% 
#   filter(Date > "1980-01-01" & Date< "2022-12-31")
# ```

#Summarize your data by month and year for graphing
DESCRIBE timestep(s) of your data here: every several years
In the example below we are averaging over sites for each month
```{r}
W6_FF <- finalFF %>%
  group_by(Year, Horizon, Elevation_grp) %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)),
            .groups = "drop")

#save these?? Right now no, and there is no data folder in the project
#write.csv(data_name, here("data", "data_name.csv"))
```

#Graph your analytes by month and year in ggplot
In this example, we have DIC and DOC
```{r}
FF_year <- ggplot(W6_FF, aes(x = Year, y = thick, col=Elevation_grp)) +
  geom_point(size = 1) +
  #geom_line() +
  labs(
    x = "Year",
    y = "Forest Floor (cm)",
    title = "FF Thickness Over Time") +
  facet_wrap(~Horizon)+
  theme_minimal()

DOC_month <- ggplot(W6_streamconc_month, aes(x = month, y = DOC)) +
  geom_point(size = 1) +
  geom_line() +
  labs(
    x = "Month",
    y = "Mean DOC (umol/L)",
    title = "Monthly Mean DOC Over Time") +
  theme_minimal()

DIC_year <- ggplot(W6_streamconc_year, aes(x = year, y = DIC)) +
  geom_point(size = 2) +
  geom_line() +
  labs(
    x = "Year",
    y = "Mean DIC (umol/L)",
    title = "Yearly Mean DIC Over Time") +
  theme_minimal()

DOC_year <- ggplot(W6_streamconc_year, aes(x = year, y = DOC)) +
  geom_point(size = 2) +
  geom_line() +
  labs(
    x = "Year",
    y = "Mean DOC (umol/L)",
    title = "Yearly Mean DOC Over Time") +
  theme_minimal()
```

#Save all your plots to the graph folder
```{r}
plots <- list(DIC_month, DOC_month, DIC_year, DOC_year) #change to your graph names
names <- c("DIC_month", "DOC_month", "DIC_year", "DOC_year") #change to names

for (i in seq_along(plots)) {
  ggsave(
    filename = here::here("Graphs", paste0(names[i], ".png")),
    plot = plots[[i]],
    width = 8,
    height = 5,
    dpi = 300
  )
}
```

LASTLY push your changes to our github repo!!